### 2025-04-08 presentations/01/main.tex
template and 01

\documentclass[aspectratio=169]{beamer} %format: 16:9

% Alternatively, select ngerman
\usepackage[american]{babel}
%\usepackage[ngerman]{babel}

% For beginners: themes need to be in the same directory as this tex-file. For experts: no explanation required. ;)
\usetheme{ITD1}
% Choose whether you want the logo on each frame title. Can be changed at any
% time. Note: It adds some vertical white space to the frame's content pane.
\withoutframelogo
%\withoutframelogo

\usepackage{csquotes}
\usepackage{derivative}
\usepackage[
  backend=biber,
  doi=true,
  eprint=false,
  date=iso,
  seconds=true,
  style=alphabetic-verb,
  locallabelwidth=true,
  maxnames = 99,
  %citestyle=alphabetic-verb
  ]{biblatex}
\renewcommand*{\bibfont}{\footnotesize}

\addbibresource{literature.bib}
%\addbibresource{jrcisia-related.bib}

%\addglobalbib{strings-short.bib}
%\addbibresource[label=jrcisia-published]{jrcisia-published.bib}

\title{An Experimental Analysis of the Influence of the Precision of Parameters in an Artificial Neural Network}
% Uncomment if unwanted
\subtitle{Natural Computation PS Presentation 01}

% \author[J. Doe, J. Doe]{Jane Doe, John Doe}
\author[D. K., E. R., S. A. S., L. S.]{Denise Katritschenko, Elias Reich, Sayed Abozar Sadat, Lukas Schwaiger}
% \institute[\isiashort, \fhshort]{\isialong\\ \depitlong\\ \fhlong}
% \institute[\fhshort]{\isialong\\ \depitlong\\ \fhlong}
\institute[\plusshort]{\pluslong\\ Department of Artificial Intelligence and Human Interfaces (AIHI)}
\date[\today]{\today}

% Automatically add fullframe for section headers. Uncomment if you do not want
% them.
\AtBeginSection[]{\fullframe{\insertsectionhead}}

\begin{document}

\frame{\titlepage}

% \section{Introduction}

% below the presentation content starts
\begin{frame}{Primer on Neural Networks}
  \begin{figure}
    \includegraphics[width=0.9\textwidth]{figures/mlp.png}
  \end{figure}
\end{frame}

\begin{frame}{Single neuron view}
  \begin{figure}
    \includegraphics[width=0.96\textwidth]{figures/single-neuron.pdf}
  \end{figure}
\end{frame}

\begin{frame}{Gradients}
  \begin{itemize}
    \item Want to optimize the parameters $\omega$ and $b$.
    \item Need a target $y$ and error metric (loss $L$) to judge the networks predictions $\hat y$.
    \item Convex optimization is used to adjust the parameters.
    \item Many intermediate results need to be stored for efficient gradient computation.
  \end{itemize}

  \begin{align*}
    L(f(x_1, x_2)) &= (\sigma(\omega^T x + b) - y) ^2\\
                   &= (\sigma(a) - y) ^ 2\\
                   &= (\hat y - y) ^ 2
  \end{align*}
  \begin{alignat*}{3}
    \pdv{L}{\omega_1} &= \pdv{L}{\sigma} &&\pdv{\sigma}{a}      &&\pdv{a}{\omega_1}\\
                      &= 2(\hat y - y)~~ &&\odv{}{a}\sigma(a)~~ &&x_1
  \end{alignat*}

\end{frame}

\begin{frame}{Memory requirements}
  For NN training we need to store the following in memory:
  \begin{itemize}
    \item Inputs and output targets (possibly batched)
    \item Parameters (weights, biases, batch statistics, ...)
    \item Intermediate results
    \item Loss
    \item Gradients
  \end{itemize}

  For inference we only need the parameters.

\end{frame}

\begin{frame}{Example}
  \begin{figure}
    \includegraphics[width=0.8\textwidth]{figures/mobilenet-size.png}
  \end{figure}
\end{frame}

\begin{frame}{Reducing memory requirements}
  Without sacrificing precision one could reduce:
  \begin{itemize}
    \item Neurons per layer
    \item Number of layers
    \item Batch size
  \end{itemize}
  Naive downsizing can quickly lead to worse generalization and expressiveness or accuracy.\\

  $\rightarrow$ Keep the number of layers and parameters and only change the
  'bitwidth' of the parameters.
\end{frame}

\begin{frame}{How parameters are represented in memory}
\centering
\includegraphics[width=0.8\textwidth]{figures/float.png}
\centering
\includegraphics[width=0.8\textwidth]{figures/binary-formats.png}
\end{frame}

\begin{frame}{Related Work}
  \begin{enumerate}
    \item Training time precision reduction\begin{itemize}
      \item Mixed Precision Training. \emph{P. Micikevicius, 2017}
      \item Training Deep Neural Network in Limited Precision. \emph{H. Park, 2018}
      \item Dorefa-Net: Training Low Bitwidth Convolutional Neural Networks with low Bitwidth Gradients. \emph{S. Zhou, 2018}
      \item Training Deep Neural Networks with 8-bit Floating Point Numbers. \emph{N. Wang, 2018}
      \item Q-BERT: Hessian-Based Ultra Low-Precision Quantization of BERT. \emph{S. Shen, 2020}

    \end{itemize}
    \item Post training precision reduction\begin{itemize}
      \item Quantization and Training of Neural Networks for Efficient Integer-Arithmetic-Only Inference. \emph{J. Benoit, 2018}
      \item GPTQ: Accurate Post-Training Quantization for Generative Pre-trained Transformers. \emph{E. Frantar, 2022}
      \item SmoothQuant: Accurate and Efficient Post-Training Quantization for Large Language Models. \emph{G. Xiao, 2024}
    \end{itemize}
  \end{enumerate}

  There are also hardware specific optimizations, e.g. FPGA \emph{P. Vera 2020}, and many more.
\end{frame}

\begin{frame}{Results from literature}
  \begin{figure}
    \includegraphics[width=0.5\textwidth]{figures/results-table.png}
    \caption{S. Zhou, 2018}
  \end{figure}
\end{frame}

\end{document}

### 2025-04-08 presentations/template/main.tex
template and 01

--- 2025-04-08 presentations/01/main.tex
@@
 
-% Alternatively, select ngerman
 \usepackage[american]{babel}
-%\usepackage[ngerman]{babel}
 
-% For beginners: themes need to be in the same directory as this tex-file. For experts: no explanation required. ;)
 \usetheme{ITD1}
-% Choose whether you want the logo on each frame title. Can be changed at any
-% time. Note: It adds some vertical white space to the frame's content pane.
 \withoutframelogo
-%\withoutframelogo
 
@@
 \addbibresource{literature.bib}
-%\addbibresource{jrcisia-related.bib}
-
-%\addglobalbib{strings-short.bib}
-%\addbibresource[label=jrcisia-published]{jrcisia-published.bib}
 
 \title{An Experimental Analysis of the Influence of the Precision of Parameters in an Artificial Neural Network}
-% Uncomment if unwanted
+
+% CHANGE IDX HERE
 \subtitle{Natural Computation PS Presentation 01}
 
-% \author[J. Doe, J. Doe]{Jane Doe, John Doe}
 \author[D. K., E. R., S. A. S., L. S.]{Denise Katritschenko, Elias Reich, Sayed Abozar Sadat, Lukas Schwaiger}
-% \institute[\isiashort, \fhshort]{\isialong\\ \depitlong\\ \fhlong}
-% \institute[\fhshort]{\isialong\\ \depitlong\\ \fhlong}
 \institute[\plusshort]{\pluslong\\ Department of Artificial Intelligence and Human Interfaces (AIHI)}
@@
 
-% Automatically add fullframe for section headers. Uncomment if you do not want
-% them.
 \AtBeginSection[]{\fullframe{\insertsectionhead}}
@@
 
-% \section{Introduction}
+\section{Introduction}
 
-% below the presentation content starts
+% each frame is a 'slide'
 \begin{frame}{Primer on Neural Networks}
   \begin{figure}
-    \includegraphics[width=0.9\textwidth]{figures/mlp.png}
+    \includegraphics[width=0.7\textwidth]{figures/mlp.png}
+    \caption{Example png.}
   \end{figure}
@@
     \includegraphics[width=0.96\textwidth]{figures/single-neuron.pdf}
+    \caption{Example pdf.}
   \end{figure}
+
+  You can still add more here.
 \end{frame}
@@
 
-\begin{frame}{Memory requirements}
-  For NN training we need to store the following in memory:
-  \begin{itemize}
-    \item Inputs and output targets (possibly batched)
-    \item Parameters (weights, biases, batch statistics, ...)
-    \item Intermediate results
-    \item Loss
-    \item Gradients
-  \end{itemize}
+\begin{frame}
+  Frame without title.\\
 
-  For inference we only need the parameters.
-
+  Example citation~\cite{Russel2020}.
 \end{frame}
 
-\begin{frame}{Example}
-  \begin{figure}
-    \includegraphics[width=0.8\textwidth]{figures/mobilenet-size.png}
-  \end{figure}
+\begin{skipframecount}
+
+\begin{frame}[allowframebreaks]{References}
+  \printbibliography[heading=none]
 \end{frame}
 
-\begin{frame}{Reducing memory requirements}
-  Without sacrificing precision one could reduce:
-  \begin{itemize}
-    \item Neurons per layer
-    \item Number of layers
-    \item Batch size
-  \end{itemize}
-  Naive downsizing can quickly lead to worse generalization and expressiveness or accuracy.\\
+\appendix
 
-  $\rightarrow$ Keep the number of layers and parameters and only change the
-  'bitwidth' of the parameters.
-\end{frame}
-
-\begin{frame}{How parameters are represented in memory}
-\centering
-\includegraphics[width=0.8\textwidth]{figures/float.png}
-\centering
-\includegraphics[width=0.8\textwidth]{figures/binary-formats.png}
-\end{frame}
-
-\begin{frame}{Related Work}
-  \begin{enumerate}
-    \item Training time precision reduction\begin{itemize}
-      \item Mixed Precision Training. \emph{P. Micikevicius, 2017}
-      \item Training Deep Neural Network in Limited Precision. \emph{H. Park, 2018}
-      \item Dorefa-Net: Training Low Bitwidth Convolutional Neural Networks with low Bitwidth Gradients. \emph{S. Zhou, 2018}
-      \item Training Deep Neural Networks with 8-bit Floating Point Numbers. \emph{N. Wang, 2018}
-      \item Q-BERT: Hessian-Based Ultra Low-Precision Quantization of BERT. \emph{S. Shen, 2020}
-
-    \end{itemize}
-    \item Post training precision reduction\begin{itemize}
-      \item Quantization and Training of Neural Networks for Efficient Integer-Arithmetic-Only Inference. \emph{J. Benoit, 2018}
-      \item GPTQ: Accurate Post-Training Quantization for Generative Pre-trained Transformers. \emph{E. Frantar, 2022}
-      \item SmoothQuant: Accurate and Efficient Post-Training Quantization for Large Language Models. \emph{G. Xiao, 2024}
-    \end{itemize}
-  \end{enumerate}
-
-  There are also hardware specific optimizations, e.g. FPGA \emph{P. Vera 2020}, and many more.
-\end{frame}
-
-\begin{frame}{Results from literature}
-  \begin{figure}
-    \includegraphics[width=0.5\textwidth]{figures/results-table.png}
-    \caption{S. Zhou, 2018}
-  \end{figure}
-\end{frame}
+\end{skipframecount}
 

### 2025-04-08 presentations/02/main.tex
presentation 02

--- 2025-04-08 presentations/template/main.tex
@@
 
-\section{Introduction}
-
-% each frame is a 'slide'
-\begin{frame}{Primer on Neural Networks}
-  \begin{figure}
-    \includegraphics[width=0.7\textwidth]{figures/mlp.png}
-    \caption{Example png.}
-  \end{figure}
+\begin{frame}{Recap and Today}
+  \textbf{What are we doing:} Change parameter precision of ANN.\\
+  \vspace{0.5cm}
+  \textbf{This presentation:} Look at the effects of
+  \begin{enumerate}
+    \item lower precision in training and inference.
+    \item lower precision only in inference.
+  \end{enumerate}
 \end{frame}
 
-\begin{frame}{Single neuron view}
-  \begin{figure}
-    \includegraphics[width=0.96\textwidth]{figures/single-neuron.pdf}
-    \caption{Example pdf.}
-  \end{figure}
-
-  You can still add more here.
-\end{frame}
-
-\begin{frame}{Gradients}
+\begin{frame}{Recap and Today}
   \begin{itemize}
     \item Want to optimize the parameters $\omega$ and $b$.
-    \item Need a target $y$ and error metric (loss $L$) to judge the networks predictions $\hat y$.
-    \item Convex optimization is used to adjust the parameters.
+    \item Need a target $y$ and error metric (loss $L$) to judge the networks predictions
+          $\hat y$.
     \item Many intermediate results need to be stored for efficient gradient computation.
@@
   \begin{align*}
-    L(f(x_1, x_2)) &= (\sigma(\omega^T x + b) - y) ^2\\
-                   &= (\sigma(a) - y) ^ 2\\
-                   &= (\hat y - y) ^ 2
+    L(f(x_1, x_2)) & = (\sigma(\omega^T x + b) - y) ^2 \\
+                   & = (\sigma(a) - y) ^ 2             \\
+                   & = (\hat y - y) ^ 2
   \end{align*}
   \begin{alignat*}{3}
-    \pdv{L}{\omega_1} &= \pdv{L}{\sigma} &&\pdv{\sigma}{a}      &&\pdv{a}{\omega_1}\\
-                      &= 2(\hat y - y)~~ &&\odv{}{a}\sigma(a)~~ &&x_1
+    \pdv{L}{\omega_1} & = \pdv{L}{\sigma} & \cdot & \pdv{\sigma}{a}      & \cdot & \pdv{a}{\omega_1} \\
+                      & = 2(\hat y - y)~~ & \cdot & \odv{}{a}\sigma(a)~~ & \cdot & x_1
   \end{alignat*}
-
 \end{frame}
 
-\begin{frame}
-  Frame without title.\\
+\begin{frame}{Data and Network Structure}
+  \textbf{MNIST:} $28 \times 28$ pixel images of handwritten digits.
 
-  Example citation~\cite{Russel2020}.
+  \begin{figure}
+    \includegraphics[width=0.7\textwidth]{figures/mnist.png}
+    \caption{MNIST dataset example images~\cite{mnistImage}.}
+  \end{figure}
 \end{frame}
 
+\begin{frame}{Data and Network Structure}
+  Model evaluation is done with with 5-fold CV.\\
+
+  \begin{figure}
+    \includegraphics[width=0.8\textwidth]{figures/mlp.png}
+    \caption{Structure of simple NN to classify MNIST numbers.}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Results - Training and Inference}
+  \begin{figure}
+    \includegraphics[width=0.7\textwidth]{figures/trainingAndInference.png}
+    \caption{Model accuracies - trained and evaluated at each precision.}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Results - Training and Inference Times}
+  \begin{figure}
+    \includegraphics[width=0.7\textwidth]{figures/times.png}
+    \caption{Training and evaluation times.}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Results - Inference-Only Testing}
+  \begin{figure}
+    \includegraphics[width=0.7\textwidth]{figures/inferenceOnly.png}
+    \caption{Model accuracies - trained with float64 and post-training precision reduction.}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Conclusion and Outlook}
+  \textbf{Interpretation:} Simple task and simple network $\rightarrow$ almost no effets.\\
+  \vspace{0.5cm}
+  \textbf{Next Steps:}
+  \begin{enumerate}
+    \item Reduce precision even further $\rightarrow$ quantize lower than float16 (half
+          precision).
+    \item Mixed Precision Training $\rightarrow$ different bit-widths for different parts
+          of the training process.
+  \end{enumerate}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%
+
 \begin{skipframecount}
-
   \begin{frame}[allowframebreaks]{References}
@@
   \appendix
-
 \end{skipframecount}

### 2025-05-28 paper/main.tex
paper init

\documentclass[11pt]{article}

% Packages
\usepackage[utf8]{inputenc}
\usepackage{amsmath, amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{titlesec}

% Geometry settings
\geometry{
	a4paper,
	total={6in, 9in},
	left=1.25in,
	top=1in,
}

% Title settings
\titleformat{\section}{\normalfont\Large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\normalfont\large\bfseries}{\thesubsection}{1em}{}

% Title
\title{An Experimental Analysis of the Influence of the Precision of
	Parameters in an Artificial Neural Network}
\author{Denise Katritschenko, Elias Reich, Sayed Abozar Sadat, Lukas Schwaiger\\
	\small University of Salzburg\\	
	\small Department of Artificial Intelligence and Human Interfaces (AIHI) }

\date{\today}

\begin{document}

	\maketitle

	\begin{abstract}
		TODO
	\end{abstract}

	\section{Introduction}
	Artificial Neural Networks (ANNs) are a key part of today’s machine learning systems. They’ve made huge progress in areas like image recognition, speech processing, and self-driving cars. However, one of the biggest challenges with ANNs is that they use a lot of memory and computing power. This becomes a problem when you want to run them on devices with limited resources, such as smartphones, embedded systems, or edge devices. \\ \\
	To solve this, researchers have looked into ways to make neural networks more efficient. A common approach is to reduce the size of the network or shrink the input batch size. But doing this can harm the network’s ability to learn and generalize. A better strategy is to reduce the numerical precision of the network’s parameters such as things like weights, activations, and gradients. This method, called quantization, makes the model smaller and faster while using less energy, without necessarily hurting accuracy. \\ \\
	One advanced version of this method is called mixed-precision quantization. Instead of using the same number of bits for every layer in the network, mixed-precision assigns different bitwidths depending on how important each part is. For example, the early layers that handle low-level features might use higher precision, while deeper layers could work fine with fewer bits. This flexible approach often leads to better results than using a single fixed precision for the whole model. Researchers have used a range of techniques, like gradient descent, heuristics, and even evolutionary algorithms, to find the best bitwidth settings for each layer. \\ \\
	Another method that helps reduce resource usage is network pruning. This means cutting out parts of the network that don’t contribute much to its performance. One interesting way to do this is with simulated annealing, a search algorithm that mimics the process of cooling metal to find a stable structure. It doesn’t need gradient information and can still find good network shapes by exploring many possibilities. This makes it useful when you want to shrink a network without going through heavy retraining. \\ \\
	Previous studies have shown that reducing the precision of neural network parameters can lead to significant improvements in memory usage and processing speed. Even when precision is reduced to very low levels, models can still perform competitively on complex tasks. However, while these early results are promising, the overall effects of lowering precision, especially how it influences model accuracy, training behavior, and storage efficiency, are still not fully understood and deserve more detailed investigation. \\ \\
	This is where our experimental work begins. 
	In this paper, we take a closer look at how different levels of parameter precision affect the behavior and performance of neural networks. Our goal is to understand what happens when we reduce precision in various ways, and how much we can lower it before performance starts to drop. \\ \\
	We start by testing how reducing precision during both training and inference affects model accuracy and runtime. Next, we look at what happens when we train the network using full precision, but lower the precision only during inference. This setup is common in real-world applications, where a powerful machine handles training, but the final model is deployed on a lightweight device. 
	Then, we run a series of tests where we round the trained weights to a limited number of decimal places. This helps us see how much we can simplify the stored weight values without hurting accuracy too much. This is especially useful for devices with very limited storage. 
	We also try out mixed-precision training, using a mix of FP16 and FP32 formats. On top of that, we experiment with different types of gradient clipping, like global norm clipping and value clipping, to see how they affect training stability and final accuracy. 
	Even though we use relatively simple network architectures and datasets, our results show that neural networks are quite robust to low-precision settings. In some cases, we were able to round weights to just three decimal digits and still get nearly the same accuracy as the full-precision version. 
	These findings show that it’s possible to significantly reduce the memory and compute needs of neural networks without losing much performance. This makes them more practical for use in low-resource settings. It also opens the door for more advanced techniques that combine quantization, pruning, and precision-aware training to build highly efficient models for real-world applications.

	\bibliographystyle{plainnat}
	\bibliography{references}

\end{document}

### 2025-05-28 paper/main.tex
first paper draft

@@
 \usepackage{geometry}
-\usepackage{natbib}
 \usepackage{setspace}
 \usepackage{titlesec}
+\usepackage{booktabs}
+\usepackage{tabularx}
+\usepackage{caption}
+\usepackage{csquotes}
+\usepackage{float}
+\usepackage{derivative}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{hyperref}
+
+\usepackage[style=nejm,
+citestyle=numeric-comp,
+sorting=none,
+doi=false,
+backend=bibtex]{biblatex}
+\addbibresource{literature.bib}
 
@@
 
-	\bibliographystyle{plainnat}
-	\bibliography{references}
+	\section{Background}
 
+	A typical neuron in a neural network performs a weighted sum of its inputs, followed by an activation function:
+	\[
+	a = \omega^\top x + b,\quad \hat{y} = \sigma(a),
+	\]
+	where \( \omega \in \mathbb{R}^n \) is the weight vector, \( b \in \mathbb{R} \) is the bias, \( x \in \mathbb{R}^n \) is the input, and \( \sigma \) is a nonlinear activation function (e.g., sigmoid or ReLU).
+
+	Training is conducted via backpropagation, where gradients of a loss function \( L(\hat{y}, y) \) with respect to the model parameters are computed. For example, using mean squared error:
+	% \[
+	% L(\hat{y}, y) = (\hat{y} - y)^2, \qquad \pdv{L}{\omega_i} = 2(\hat{y} - y) \cdot \odv{\sigma}{a} \cdot x_i.
+	% \]
+
+  	\begin{align*}
+    	L(f(x_1, x_2)) &= (\sigma(\omega^T x + b) - y) ^2\\
+                   	&= (\sigma(a) - y) ^ 2\\
+                   	&= (\hat y - y) ^ 2
+  	\end{align*}
+  	\begin{alignat*}{3}
+    	\pdv{L}{\omega_1} &= \pdv{L}{\sigma} &&\pdv{\sigma}{a}      &&\pdv{a}{\omega_1}\\
+                      	&= 2(\hat y - y)~~ &&\odv{}{a}\sigma(a)~~ &&x_1
+  	\end{alignat*}
+
+	Memory and compute requirements during training are substantial, as gradients, activations, and parameters must be stored and updated iteratively.
+In contrast, inference requires only a forward pass, making it more amenable to optimization via precision reduction.
+
+	\section{Memory Efficiency and Numerical Formats}
+
+	To reduce memory usage and computational load, precision-aware techniques such as quantization and mixed-precision training are increasingly employed.
+Instead of using 32-bit or 64-bit floating-point formats, parameters can be stored in 16-bit or even 8-bit fixed-point representations.
+
+	\begin{figure}[H]
+    	\centering
+    	\includegraphics[width=0.8\textwidth]{figures/float.png}
+    	\caption{Illustration of standard floating-point format.}
+	\end{figure}
+
+	\begin{figure}[H]
+    	\centering
+    	\includegraphics[width=0.8\textwidth]{figures/binary-formats.png}
+    	\caption{Comparison of binary formats used in neural network computations.}
+	\end{figure}
+
+	\section{Related Work}
+
+	Prior research in this domain can be broadly classified into two categories:
+
+	\subsection*{Quantization During Training}
+	Several studies have explored training neural networks with reduced precision directly. Notable contributions include:
+	\begin{itemize}
+    	\item \textbf{Mixed Precision Training}, which leverages float16 arithmetic for speed while maintaining float32 master weights \cite{micikevicius2017mixed}.
+    	\item \textbf{DoReFa-Net}, which quantizes weights, activations, and gradients to arbitrary bitwidths \cite{zhou2018dorefa}.
+    	\item \textbf{Q-BERT}, a quantized BERT model that maintains accuracy using only 4-bit weights \cite{shen2020qbert}.
+	\end{itemize}
+
+	\subsection*{Post-Training Quantization}
+	Other approaches focus on converting pretrained models into lower-precision equivalents:
+	\begin{itemize}
+    	\item \textbf{Integer-Only Quantization} using symmetric quantization scales \cite{jacob2018quantization}.
+    	\item \textbf{GPTQ}, a method for quantizing large transformer models without retraining \cite{frantar2022gptq}.
+    	\item \textbf{SmoothQuant}, which scales activations to better match weight ranges during quantization \cite{xiao2024smoothquant}.
+	\end{itemize}
+
+	\section{Experimental Design}
+
+	We trained a simple fully connected neural network on the MNIST dataset, consisting of 60,000 training and 10,000 test grayscale images of handwritten digits (28×28 pixels).
+Model training was performed under varying parameter bitwidths, with cross-validation to ensure robustness.
+
+	\begin{figure}[H]
+    	\centering
+    	\includegraphics[width=0.75\textwidth]{figures/mnist.png}
+    	\caption{Sample images from the MNIST dataset.}
+	\end{figure}
+
+	\section{Results}
+
+	\subsection*{Accuracy Across Bitwidths}
+
+	\begin{figure}[H]
+    	\centering
+    	\includegraphics[width=0.75\textwidth]{figures/trainingAndInference.png}
+    	\caption{Model accuracy during training and inference under different precision settings.}
+	\end{figure}
+
+	\subsection*{Training and Inference Time}
+
+	\begin{figure}[H]
+    	\centering
+    	\includegraphics[width=0.75\textwidth]{figures/times.png}
+    	\caption{Execution times for various precision configurations.}
+	\end{figure}
+
+	\subsection*{Quantization After Training}
+
+	\begin{figure}[H]
+    	\centering
+    	\includegraphics[width=0.75\textwidth]{figures/inferenceOnly.png}
+    	\caption{Inference accuracy using float64-trained weights followed by quantization.}
+	\end{figure}
+
+	\subsection*{Automatic Mixed Precision and Gradient Clipping}
+	\begin{figure}[H]
+    	\centering
+    	\includegraphics[width=0.75\textwidth]{figures/gradclip_32.png}
+    	\caption{Inference accuracy using float32-trained weights with AMP and gradient clipping.}
+	\end{figure}
+
+	\subsection*{Effect of Decimal Rounding}
+
+	\begin{figure}[H]
+    	\centering
+    	\includegraphics[width=0.75\textwidth]{figures/118_accuracy.png}
+    	\caption{Accuracy vs. number of decimal digits retained in weights.}
+	\end{figure}
+
+	\section{Discussion}
+
+	Our results indicate that networks trained and evaluated with as low as 16-bit precision can achieve comparable performance to their high-precision counterparts.
+Moreover, rounding pretrained weights to just 3 decimal digits preserves full accuracy, opening the door to aggressive post-training quantization.
+
+	This is particularly relevant for edge devices and real-time applications where memory and compute constraints are paramount.
+
+	\section{Conclusion and Future Directions}
+
+	This work confirms that for relatively simple tasks like MNIST classification, significant reductions in parameter precision can be achieved without compromising performance.
+Our future efforts will extend this analysis to more complex datasets (e.g., CIFAR-10, ImageNet) and deeper architectures, and explore advanced quantization-aware training techniques.
+
+	\newpage
+	\printbibliography
 \end{document}

### 2025-07-12 paper/main.tex
formatting

@@
 
-% Packages
 \usepackage[utf8]{inputenc}
@@
 \section{Introduction}
-	Artificial Neural Networks (ANNs) are a key part of today’s machine learning systems.
-They’ve made huge progress in areas like image recognition, speech processing, and self-driving cars.
-However, one of the biggest challenges with ANNs is that they use a lot of memory and computing power.
-This becomes a problem when you want to run them on devices with limited resources, such as smartphones, embedded systems, or edge devices.
-\\ \\
-	To solve this, researchers have looked into ways to make neural networks more efficient.
-A common approach is to reduce the size of the network or shrink the input batch size.
-But doing this can harm the network’s ability to learn and generalize.
-A better strategy is to reduce the numerical precision of the network’s parameters such as things like weights, activations, and gradients.
-This method, called quantization, makes the model smaller and faster while using less energy, without necessarily hurting accuracy.
-\\ \\
-	One advanced version of this method is called mixed-precision quantization.
-Instead of using the same number of bits for every layer in the network, mixed-precision assigns different bitwidths depending on how important each part is.
-For example, the early layers that handle low-level features might use higher precision, while deeper layers could work fine with fewer bits.
-This flexible approach often leads to better results than using a single fixed precision for the whole model.
-Researchers have used a range of techniques, like gradient descent, heuristics, and even evolutionary algorithms, to find the best bitwidth settings for each layer.
-\\ \\
-	Another method that helps reduce resource usage is network pruning.
-This means cutting out parts of the network that don’t contribute much to its performance.
-One interesting way to do this is with simulated annealing, a search algorithm that mimics the process of cooling metal to find a stable structure.
-It doesn’t need gradient information and can still find good network shapes by exploring many possibilities.
-This makes it useful when you want to shrink a network without going through heavy retraining.
-\\ \\
-	Previous studies have shown that reducing the precision of neural network parameters can lead to significant improvements in memory usage and processing speed.
-Even when precision is reduced to very low levels, models can still perform competitively on complex tasks.
-However, while these early results are promising, the overall effects of lowering precision, especially how it influences model accuracy, training behavior, and storage efficiency, are still not fully understood and deserve more detailed investigation.
-\\ \\
-	This is where our experimental work begins. 
-	In this paper, we take a closer look at how different levels of parameter precision affect the behavior and performance of neural networks.
-Our goal is to understand what happens when we reduce precision in various ways, and how much we can lower it before performance starts to drop.
-\\ \\
-	We start by testing how reducing precision during both training and inference affects model accuracy and runtime.
-Next, we look at what happens when we train the network using full precision, but lower the precision only during inference.
-This setup is common in real-world applications, where a powerful machine handles training, but the final model is deployed on a lightweight device.
-	Then, we run a series of tests where we round the trained weights to a limited number of decimal places.
-This helps us see how much we can simplify the stored weight values without hurting accuracy too much.
-This is especially useful for devices with very limited storage.
-	We also try out mixed-precision training, using a mix of FP16 and FP32 formats.
-On top of that, we experiment with different types of gradient clipping, like global norm clipping and value clipping, to see how they affect training stability and final accuracy.
-	Even though we use relatively simple network architectures and datasets, our results show that neural networks are quite robust to low-precision settings.
-In some cases, we were able to round weights to just three decimal digits and still get nearly the same accuracy as the full-precision version.
-	These findings show that it’s possible to significantly reduce the memory and compute needs of neural networks without losing much performance.
-This makes them more practical for use in low-resource settings.
-It also opens the door for more advanced techniques that combine quantization, pruning, and precision-aware training to build highly efficient models for real-world applications.
+Artificial Neural Networks (ANNs) are a key part of today’s machine learning
+systems. They’ve made huge progress in areas like image recognition, speech
+processing, and self-driving cars. However, one of the biggest challenges with
+ANNs is that they use a lot of memory and computing power. This becomes a
+problem when you want to run them on devices with limited resources, such as
+smartphones, embedded systems, or edge devices. \\ \\ To solve this,
+researchers have looked into ways to make neural networks more efficient. A
+common approach is to reduce the size of the network or shrink the input batch
+size. But doing this can harm the network’s ability to learn and generalize. A
+better strategy is to reduce the numerical precision of the network’s
+parameters such as things like weights, activations, and gradients. This
+method, called quantization, makes the model smaller and faster while using
+less energy, without necessarily hurting accuracy. \\ \\ One advanced version
+of this method is called mixed-precision quantization. Instead of using the
+same number of bits for every layer in the network, mixed-precision assigns
+different bitwidths depending on how important each part is. For example, the
+early layers that handle low-level features might use higher precision, while
+deeper layers could work fine with fewer bits. This flexible approach often
+leads to better results than using a single fixed precision for the whole
+model. Researchers have used a range of techniques, like gradient descent,
+heuristics, and even evolutionary algorithms, to find the best bitwidth
+settings for each layer. \\ \\ Another method that helps reduce resource usage
+is network pruning. This means cutting out parts of the network that don’t
+contribute much to its performance. One interesting way to do this is with
+simulated annealing, a search algorithm that mimics the process of cooling
+metal to find a stable structure. It doesn’t need gradient information and can
+still find good network shapes by exploring many possibilities. This makes it
+useful when you want to shrink a network without going through heavy
+retraining. \\ \\ Previous studies have shown that reducing the precision of
+neural network parameters can lead to significant improvements in memory usage
+and processing speed. Even when precision is reduced to very low levels, models
+can still perform competitively on complex tasks. However, while these early
+results are promising, the overall effects of lowering precision, especially
+how it influences model accuracy, training behavior, and storage efficiency,
+are still not fully understood and deserve more detailed investigation. \\ \\
+This is where our experimental work begins. In this paper, we take a closer
+look at how different levels of parameter precision affect the behavior and
+performance of neural networks. Our goal is to understand what happens when we
+reduce precision in various ways, and how much we can lower it before
+performance starts to drop. \\ \\ We start by testing how reducing precision
+during both training and inference affects model accuracy and runtime. Next, we
+look at what happens when we train the network using full precision, but lower
+the precision only during inference. This setup is common in real-world
+applications, where a powerful machine handles training, but the final model is
+deployed on a lightweight device. Then, we run a series of tests where we round
+the trained weights to a limited number of decimal places. This helps us see
+how much we can simplify the stored weight values without hurting accuracy too
+much. This is especially useful for devices with very limited storage. We also
+try out mixed-precision training, using a mix of FP16 and FP32 formats. On top
+of that, we experiment with different types of gradient clipping, like global
+norm clipping and value clipping, to see how they affect training stability and
+final accuracy. Even though we use relatively simple network architectures and
+datasets, our results show that neural networks are quite robust to
+low-precision settings. In some cases, we were able to round weights to just
+three decimal digits and still get nearly the same accuracy as the
+full-precision version. These findings show that it’s possible to significantly
+reduce the memory and compute needs of neural networks without losing much
+performance. This makes them more practical for use in low-resource settings.
+It also opens the door for more advanced techniques that combine quantization,
+pruning, and precision-aware training to build highly efficient models for
+real-world applications.
 
@@
 
-	A typical neuron in a neural network performs a weighted sum of its inputs, followed by an activation function:
+A typical neuron in a neural network performs a weighted sum of its inputs,
+followed by an activation function:
 \[
@@
 \]
-	where \( \omega \in \mathbb{R}^n \) is the weight vector, \( b \in \mathbb{R} \) is the bias, \( x \in \mathbb{R}^n \) is the input, and \( \sigma \) is a nonlinear activation function (e.g., sigmoid or ReLU).
+where \( \omega \in \mathbb{R}^n \) is the weight vector, \( b \in \mathbb{R}
+\) is the bias, \( x \in \mathbb{R}^n \) is the input, and \( \sigma \) is a
+nonlinear activation function (e.g., sigmoid or ReLU).
 
-	Training is conducted via backpropagation, where gradients of a loss function \( L(\hat{y}, y) \) with respect to the model parameters are computed. For example, using mean squared error:
+Training is conducted via backpropagation, where gradients of a loss function
+\( L(\hat{y}, y) \) with respect to the model parameters are computed. For
+example, using mean squared error:
 % \[
@@
 \begin{align*}
-    	L(f(x_1, x_2)) &= (\sigma(\omega^T x + b) - y) ^2\\
-                   	&= (\sigma(a) - y) ^ 2\\
-                   	&= (\hat y - y) ^ 2
+	L(f(x_1, x_2)) & = {(\sigma(\omega^T x + b) - y)}^2 \\
+	               & = {(\sigma(a) - y)}^2              \\
+	               & = {(\hat y - y)}^2
 \end{align*}
 \begin{alignat*}{3}
-    	\pdv{L}{\omega_1} &= \pdv{L}{\sigma} &&\pdv{\sigma}{a}      &&\pdv{a}{\omega_1}\\
-                      	&= 2(\hat y - y)~~ &&\odv{}{a}\sigma(a)~~ &&x_1
+	\pdv{L}{\omega_1} & = \pdv{L}{\sigma} &  & \pdv{\sigma}{a}      &  & \pdv{a}{\omega_1} \\
+	                  & = 2(\hat y - y)~~ &  & \odv{}{a}\sigma(a)~~ &  & x_1
 \end{alignat*}
 
-	Memory and compute requirements during training are substantial, as gradients, activations, and parameters must be stored and updated iteratively.
-In contrast, inference requires only a forward pass, making it more amenable to optimization via precision reduction.
+Memory and compute requirements during training are substantial, as gradients,
+activations, and parameters must be stored and updated iteratively. In
+contrast, inference requires only a forward pass, making it more amenable to
+optimization via precision reduction.
 
@@
 
-	To reduce memory usage and computational load, precision-aware techniques such as quantization and mixed-precision training are increasingly employed.
-Instead of using 32-bit or 64-bit floating-point formats, parameters can be stored in 16-bit or even 8-bit fixed-point representations.
+To reduce memory usage and computational load, precision-aware techniques such
+as quantization and mixed-precision training are increasingly employed. Instead
+of using 32-bit or 64-bit floating-point formats, parameters can be stored in
+16-bit or even 8-bit fixed-point representations.
 
@@
 \subsection*{Quantization During Training}
-	Several studies have explored training neural networks with reduced precision directly. Notable contributions include:
+Several studies have explored training neural networks with reduced precision
+directly. Notable contributions include:
 \begin{itemize}
-    	\item \textbf{Mixed Precision Training}, which leverages float16 arithmetic for speed while maintaining float32 master weights \cite{micikevicius2017mixed}.
-    	\item \textbf{DoReFa-Net}, which quantizes weights, activations, and gradients to arbitrary bitwidths \cite{zhou2018dorefa}.
-    	\item \textbf{Q-BERT}, a quantized BERT model that maintains accuracy using only 4-bit weights \cite{shen2020qbert}.
+	\item \textbf{Mixed Precision Training}, which leverages float16 arithmetic for speed while maintaining float32 master weights~\cite{micikevicius2017mixed}.
+	\item \textbf{DoReFa-Net}, which quantizes weights, activations, and gradients to arbitrary bitwidths~\cite{zhou2018dorefa}.
+	\item \textbf{Q-BERT}, a quantized BERT model that maintains accuracy using only 4-bit weights~\cite{shen2020qbert}.
 \end{itemize}
@@
 \subsection*{Post-Training Quantization}
-	Other approaches focus on converting pretrained models into lower-precision equivalents:
+Other approaches focus on converting pretrained models into lower-precision
+equivalents:
 \begin{itemize}
-    	\item \textbf{Integer-Only Quantization} using symmetric quantization scales \cite{jacob2018quantization}.
-    	\item \textbf{GPTQ}, a method for quantizing large transformer models without retraining \cite{frantar2022gptq}.
-    	\item \textbf{SmoothQuant}, which scales activations to better match weight ranges during quantization \cite{xiao2024smoothquant}.
+	\item \textbf{Integer-Only Quantization} using symmetric quantization scales~\cite{jacob2018quantization}.
+	\item \textbf{GPTQ}, a method for quantizing large transformer models without retraining~\cite{frantar2022gptq}.
+	\item \textbf{SmoothQuant}, which scales activations to better match weight ranges during quantization~\cite{xiao2024smoothquant}.
 \end{itemize}
@@
 
-	We trained a simple fully connected neural network on the MNIST dataset, consisting of 60,000 training and 10,000 test grayscale images of handwritten digits (28×28 pixels).
-Model training was performed under varying parameter bitwidths, with cross-validation to ensure robustness.
+We trained a simple fully connected neural network on the MNIST dataset,
+consisting of 60,000 training and 10,000 test grayscale images of handwritten
+digits (28×28 pixels). Model training was performed under varying parameter
+bitwidths, with cross-validation to ensure robustness.
 
@@
 	\includegraphics[width=0.75\textwidth]{figures/118_accuracy.png}
-    	\caption{Accuracy vs. number of decimal digits retained in weights.}
+	\caption{Accuracy vs.\ number of decimal digits retained in weights.}
 \end{figure}
@@
 
-	Our results indicate that networks trained and evaluated with as low as 16-bit precision can achieve comparable performance to their high-precision counterparts.
-Moreover, rounding pretrained weights to just 3 decimal digits preserves full accuracy, opening the door to aggressive post-training quantization.
+Our results indicate that networks trained and evaluated with as low as 16-bit
+precision can achieve comparable performance to their high-precision
+counterparts. Moreover, rounding pretrained weights to just 3 decimal digits
+preserves full accuracy, opening the door to aggressive post-training
+quantization.
 
-	This is particularly relevant for edge devices and real-time applications where memory and compute constraints are paramount.
+This is particularly relevant for edge devices and real-time applications where
+memory and compute constraints are paramount.
 
@@
 
-	This work confirms that for relatively simple tasks like MNIST classification, significant reductions in parameter precision can be achieved without compromising performance.
-Our future efforts will extend this analysis to more complex datasets (e.g., CIFAR-10, ImageNet) and deeper architectures, and explore advanced quantization-aware training techniques.
+This work confirms that for relatively simple tasks like MNIST classification,
+significant reductions in parameter precision can be achieved without
+compromising performance. Our future efforts will extend this analysis to more
+complex datasets (e.g., CIFAR-10, ImageNet) and deeper architectures, and
+explore advanced quantization-aware training techniques.
 
 \newpage
-	\printbibliography
+\printbibliography%
 \end{document}

### 2025-07-12 presentations/05/main.tex
praes05 upload

\documentclass[aspectratio=169]{beamer} %format: 16:9

\usepackage[american]{babel}

\usetheme{ITD1}
\withoutframelogo

\usepackage{booktabs}
\usepackage{csquotes}
\usepackage{derivative}
\usepackage[
  backend=biber,
  doi=true,
  eprint=false,
  date=iso,
  seconds=true,
  style=alphabetic-verb,
  locallabelwidth=true,
  maxnames = 99,
  %citestyle=alphabetic-verb
]{biblatex}
\renewcommand*{\bibfont}{\footnotesize}

\addbibresource{literature.bib}

\title{An Experimental Analysis of the Influence of the Precision of Parameters in an Artificial Neural Network}

% CHANGE IDX HERE
\subtitle{Natural Computation PS Presentation 05}

\author[D. K., E. R., S. A. S., L. S.]{Denise Katritschenko, Elias Reich, Sayed Abozar Sadat, Lukas Schwaiger}
\institute[\plusshort]{\pluslong\\ Department of Artificial Intelligence and Human Interfaces (AIHI)}
\date[\today]{\today}

\AtBeginSection[]{\fullframe{\insertsectionhead}}

\begin{document}

\frame{\titlepage}

\begin{frame}{DoReFa-Net: Introduction}
  \textbf{Last times:} Impact of quantizing fractional parts of the weights to various decimals, training/inference using fp64/fp32/fp16, \dots\\
  \vspace{0.5cm}
  \textbf{This presentation:}
  \begin{itemize}
    \item Quantization in \textit{all} numerical domains: Weights, Activations, and Gradients.
    \item Consideration of \textit{task complexity} $\rightarrow$ dataset and classes.
    \item \textbf{DoReFa-Net}\cite{zhou2016dorefa}: Low-bit running and training of CNN on ImageNet-scale models.
  \end{itemize}
\end{frame}

\begin{frame}{DoReFa-Net: Selected Ideas}
  \begin{itemize}
    \item \textbf{Different bit-widths:} For Weights, Activations, and Gradients for accuracy/efficiency trade-off.
    \item \textbf{Bit-convolution kernel:} Any $M\times K$-bit dot product becomes AND+popcount operation (or XNOR).
    \item \textbf{Different precisions per tensor:} Typically $W\le A \le G$ (e.g., ImageNet $1\text{W}/2\text{A}/6\text{G}$).
    \item \textbf{Straight-Through Estimator (STE):} Keep normal gradients also through non-differentiable quantizers.
  \end{itemize}
\end{frame}

\begin{frame}{DoReFa-Net: Training Pipeline}
  \begin{enumerate}
    \item \textbf{Forward pass}
          \begin{itemize}
            \item Quantize current weights $W^{full-prec}\xrightarrow[]{f_\omega}W^{bit-width}$
            \item Propagate activations, then quantize them $A^{real-valued}\xrightarrow[]{f_\alpha}A^{bit-width}$
          \end{itemize}
    \item \textbf{Backward pass}
          \begin{itemize}
            \item Back-propagate loss - each gradient block is quantized $G\xrightarrow[]{f_\gamma}G^{bit-width}$
          \end{itemize}
  \end{enumerate}
\end{frame}

\begin{frame}{DoReFa-Net: First \& Last FC Layers}
  \begin{itemize}
    \item Quantizing either layer costs $\approx 0.3$pp on SVHN Model A. Smaller models suffer more (up to $1$ pp).
    \item Paper keeps those layers full-precision (fp32) in ImageNet runs.
  \end{itemize}
\end{frame}

\begin{frame}{DoReFa-Net: Accuracy AlexNet on ImageNet}
  \begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/dorefaRes.png}
    \caption*{\scriptsize Generated bar chart of Top-1 accuracies.}
  \end{figure}
\end{frame}

\begin{frame}{DoReFa-Net: Task Complexity Matters}
  Accuracies of DoReFa-Net on \textbf{SVHN} (10-class house number digit dataset) are very high.\\
  \textbf{Model A} = Full-width 7-layer CNN, trained from scratch with DoReFa quantisation.\\
  \textbf{Model D} = $\frac{1}{8}$-width variant (87.5\% fewer channels $\rightarrow$ network keeps only 12.5\% of the original feature-map width in every convolutional layer). %Example: If Model A has 128 channels in a layer, Model D uses 128x(1−0.875)=128x0.125=16 channels
  \vspace{0.3cm}

  \renewcommand{\arraystretch}{0.85}
  \centering
  \begin{columns}[T,onlytextwidth]
    \column{0.49\textwidth}
    \begin{tabular}{ccc cc}
      \toprule
      $W$ & $A$ & $G$ & Model A & Model D \\
      \midrule
      1   & 1   & 2   & 0.934   & 0.803   \\
      1   & 1   & 4   & 0.968   & 0.846   \\
      1   & 1   & 8   & 0.970   & 0.828   \\
      1   & 1   & 32  & 0.971   & 0.841   \\
      1   & 2   & 2   & 0.909   & 0.808   \\
      1   & 2   & 3   & 0.968   & 0.878   \\
      1   & 2   & 4   & 0.975   & 0.878   \\
      1   & 2   & 8   & 0.975   & 0.866   \\
      1   & 2   & 32  & 0.976   & 0.865   \\
      2   & 1   & 2   & 0.927   & 0.846   \\
      \bottomrule
    \end{tabular}

    \column{0.49\textwidth}
    \begin{tabular}{ccc cc}
      \toprule
      $W$ & $A$ & $G$ & Model A & Model D \\
      \midrule
      2   & 1   & 4   & 0.969   & 0.827   \\
      1   & 3   & 3   & 0.968   & 0.887   \\
      1   & 3   & 4   & 0.974   & 0.897   \\
      1   & 3   & 6   & 0.977   & 0.916   \\
      1   & 4   & 2   & 0.815   & 0.868   \\
      1   & 4   & 4   & 0.975   & 0.915   \\
      1   & 4   & 8   & 0.977   & 0.895   \\
      2   & 2   & 8   & 0.900   & 0.842   \\
      8   & 8   & 8   & 0.970   & 0.955   \\
      32  & 32  & 32  & 0.975   & 0.950   \\
      \bottomrule
    \end{tabular}
  \end{columns}
\end{frame}

\begin{frame}{DoReFa-Net: Learnings \& Outlook}
  \begin{itemize}
    \item Gradient precision (min. 6 bits) is the bottleneck - mixed precision was the next step.
    \item Quantization is not \textit{just for deployment} - it can be applied inside the training pipeline.
  \end{itemize}
\end{frame}

% \begin{frame}{DoReFa-Net: Introduction}
%   \textbf{What are we doing:} Change parameter precision of ANN.\\
%   \vspace{0.5cm}
%   \textbf{This presentation:} Look at the effects of
%   \begin{enumerate}
%     \item lower precision in training and inference.
%     \item lower precision only in inference.
%   \end{enumerate}
% \end{frame}

% \begin{frame}{DoReFa-Net: Network Structure}
%   \textbf{MNIST:} $28 \times 28$ pixel images of handwritten digits.

%   \begin{figure}
%     \includegraphics[width=0.7\textwidth]{figures/mnist.png}
%     \caption{MNIST dataset example images~\cite{mnistImage}.}
%   \end{figure}
% \end{frame}

% \begin{frame}{Recap and Today}
%   \begin{itemize}
%     \item Want to optimize the parameters $\omega$ and $b$.
%     \item Need a target $y$ and error metric (loss $L$) to judge the networks predictions
%           $\hat y$.
%     \item Many intermediate results need to be stored for efficient gradient computation.
%   \end{itemize}

%   \begin{align*}
%     L(f(x_1, x_2)) & = (\sigma(\omega^T x + b) - y) ^2 \\
%                    & = (\sigma(a) - y) ^ 2             \\
%                    & = (\hat y - y) ^ 2
%   \end{align*}
%   \begin{alignat*}{3}
%     \pdv{L}{\omega_1} & = \pdv{L}{\sigma} & \cdot & \pdv{\sigma}{a}      & \cdot & \pdv{a}{\omega_1} \\
%                       & = 2(\hat y - y)~~ & \cdot & \odv{}{a}\sigma(a)~~ & \cdot & x_1
%   \end{alignat*}
% \end{frame}

% \begin{frame}{Data and Network Structure}
%   Model evaluation is done with with 5-fold CV.\\

%   \begin{figure}
%     \includegraphics[width=0.8\textwidth]{figures/mlp.png}
%     \caption{Structure of simple NN to classify MNIST numbers.}
%   \end{figure}
% \end{frame}

% \begin{frame}{Results - Training and Inference}
%   \begin{figure}
%     \includegraphics[width=0.7\textwidth]{figures/trainingAndInference.png}
%     \caption{Model accuracies - trained and evaluated at each precision.}
%   \end{figure}
% \end{frame}

% \begin{frame}{Results - Training and Inference Times}
%   \begin{figure}
%     \includegraphics[width=0.7\textwidth]{figures/times.png}
%     \caption{Training and evaluation times.}
%   \end{figure}
% \end{frame}

% \begin{frame}{Results - Inference-Only Testing}
%   \begin{figure}
%     \includegraphics[width=0.7\textwidth]{figures/inferenceOnly.png}
%     \caption{Model accuracies - trained with float64 and post-training precision reduction.}
%   \end{figure}
% \end{frame}

% \begin{frame}{Conclusion and Outlook}
%   \textbf{Interpretation:} Simple task and simple network $\rightarrow$ almost no effets.\\
%   \vspace{0.5cm}
%   \textbf{Next Steps:}
%   \begin{enumerate}
%     \item Reduce precision even further $\rightarrow$ quantize lower than float16 (half
%           precision).
%     \item Mixed Precision Training $\rightarrow$ different bit-widths for different parts
%           of the training process.
%   \end{enumerate}
% \end{frame}

%%%%%%%%%%%%%%%%%%%%

\begin{skipframecount}
  \begin{frame}[allowframebreaks]{References}
    \printbibliography[heading=none]
  \end{frame}

  \appendix
\end{skipframecount}

\end{document}

### 2025-07-15 paper/main.tex
praes05 upload
kurz vor Klausur

@@
 \usepackage{graphicx}
-\usepackage{hyperref}
+\usepackage[hidelinks]{hyperref}
 \usepackage{geometry}
@@
 \usepackage{derivative}
-\usepackage{caption}
 \usepackage{subcaption}
-\usepackage{hyperref}
 
-\usepackage[style=nejm,
+\usepackage[
+	style=nejm,
 	citestyle=numeric-comp,
@@
 % Title
-\title{An Experimental Analysis of the Influence of the Precision of
-	Parameters in an Artificial Neural Network}
+\title{An Experimental Analysis of the Influence of the Precision of Parameters in an Artificial Neural Network}
 \author{Denise Katritschenko, Elias Reich, Sayed Abozar Sadat, Lukas Schwaiger\\
@@
 \begin{abstract}
-	TODO
+	We study how the numerical precision of weights, activations, and gradients in neural networks can be reduced without sacrificing too much of the classification performance.
+Using MNIST as a dataset, we benchmark uniform 16–64-bit quantisation, mixed-precision formats, and post-training decimal rounding.
+The results reveal that (i) 16-bit end-to-end training preserves full accuracy for the MNIST dataset and the network used, (ii) three-decimal-place weight rounding while keeping 32-bit precision incurs minimal accuracy loss.
+These findings verify limits for memory- and energy-constrained usecases like inference in mobile devices.
 \end{abstract}
@@
 \( L(\hat{y}, y) \) with respect to the model parameters are computed. For
-example, using mean squared error:
+example, using the mean squared error:
 % \[
@@
 
-Memory and compute requirements during training are substantial, as gradients,
-activations, and parameters must be stored and updated iteratively. In
-contrast, inference requires only a forward pass, making it more amenable to
+Memory and computational requirements during training are substantial, as
+gradients, activations, and parameters must be stored and updated iteratively.
+In contrast, inference requires only a forward pass, making it more amenable to
 optimization via precision reduction.
@@
 16-bit or even 8-bit fixed-point representations.
+Figure~\ref{fig:floatingPointFormat} depicts the general floating point format
+and terminology of significand, base, and exponent.
+Tables~\ref{tab:binary-bits} and~\ref{tab:binary-precision} provide an overview
+of the different floating point types and their respective characteristics.
 
@@
 	\includegraphics[width=0.8\textwidth]{figures/float.png}
-	\caption{Illustration of standard floating-point format.}
+	\caption{Illustration of standard floating-point format.}\label{fig:floatingPointFormat}
 \end{figure}
 
-\begin{figure}[H]
+\begin{table}[H]
 	\centering
-	\includegraphics[width=0.8\textwidth]{figures/binary-formats.png}
-	\caption{Comparison of binary formats used in neural network computations.}
-\end{figure}
+	\begin{tabular}{@{}lcccc@{}}
+		\toprule
+		\textbf{Type} & Sign & Exponent & Significand & Total \\
+		\midrule
+		Half  (FP16)  & 1    & 5        & 10          & 16    \\
+		Single (FP32) & 1    & 8        & 23          & 32    \\
+		Double (FP64) & 1    & 11       & 52          & 64    \\
+		\bottomrule
+	\end{tabular}
+	\caption{Field widths (in bits) of common IEEE754 formats (adapted from~\cite{fparithmetic}).}\label{tab:binary-bits}
+\end{table}
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}{@{}lccc@{}}
+		\toprule
+		\textbf{Type} & Exponent bias & Precision (bits) & Decimal digits \\
+		\midrule
+		Half  (FP16)  & 15            & 11               & $\sim$3.3      \\
+		Single (FP32) & 127           & 24               & $\sim$7.2      \\
+		Double (FP64) & 1023          & 53               & $\sim$15.9     \\
+		\bottomrule
+	\end{tabular}
+	\caption{Range and precision characteristics of the same formats (adapted from~\cite{fparithmetic}).}\label{tab:binary-precision}
+\end{table}
 
@@
 consisting of 60,000 training and 10,000 test grayscale images of handwritten
-digits (28×28 pixels). Model training was performed under varying parameter
-bitwidths, with cross-validation to ensure robustness.
+digits (28×28 pixels, see figure~\ref{fig:mnistSample}). Model training was
+performed under varying parameter bitwidths, and 5-fold cross-validation to
+evaluate the model performance. The structure of the MLP is depicted in
+figure~\ref{fig:mlpStructure}.
 
@@
 	\centering
-	\includegraphics[width=0.75\textwidth]{figures/mnist.png}
-	\caption{Sample images from the MNIST dataset.}
+	\includegraphics[width=0.85\textwidth]{figures/mnist.png}
+	\caption{Sample images from the MNIST dataset.}\label{fig:mnistSample}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.85\textwidth]{figures/mlpStruc.png}
+	\caption{MLP structure used in the experiments.}\label{fig:mlpStructure}
 \end{figure}
@@
 \section{Results}
+In the first experiment, every floating-point tensor that is used inside the
+training loop was converted to a lower precision data type. This therefore
+includes the weights and biases, activations, gradients, optimizer states, etc.
+The types tried were float64, float32, and float16.
 
 \subsection*{Accuracy Across Bitwidths}
-
+Figure~\ref{fig:modelAccTrainAndInf} shows how the accuracy changed after the
+models were trained on the different precision settings. Keep in mind that the
+y-axis is not starting from zero, so the absolute height differences are
+amplified visually, though the delta is only around $.1$ percent. Re-training
+the models and re-evaluating them has also lead to the accuracies being
+``reversed'', meaning the float64 model had the least accuracy. It is safe to
+assume that the model accuracy has neither decreased, nor increased.
 \begin{figure}[H]
 	\centering
-	\includegraphics[width=0.75\textwidth]{figures/trainingAndInference.png}
-	\caption{Model accuracy during training and inference under different precision settings.}
+	\includegraphics[width=0.85\textwidth]{figures/trainingAndInference.png}
+	\caption{Model accuracy evaluated after training with the different precisions.}\label{fig:modelAccTrainAndInf}
 \end{figure}
@@
 \subsection*{Training and Inference Time}
-
+Also analyzed were the times needed for both the training and inference.
+Figure~\ref{fig:modelEvalTrainTimes} depicts the times measured in the training
+loop and later on at the inference stage. Again, the times are almost identical
+in both regards, and neither a real speed-up nor a slow-down has been measured.
 \begin{figure}[H]
 	\centering
-	\includegraphics[width=0.75\textwidth]{figures/times.png}
-	\caption{Execution times for various precision configurations.}
+	\includegraphics[width=0.85\textwidth]{figures/times.png}
+	\caption{Training and evaluation times for the different precisions.}\label{fig:modelEvalTrainTimes}
 \end{figure}
 
-\subsection*{Quantization After Training}
-
+\subsection*{Inference-Only Testing}
+This third experiment was conducted to test a possible post-training precision
+reduction. For this, the model was trained on the highest precision (float64)
+and then quantized at the inference stage.
+Figure~\ref{fig:inferenceOnlyTesting} shows that the accuracy of the baseline
+(float64 training AND inference) is maintained as the precision is decreased at
+the later stages. In contrast to the \textit{Accuracy Across Bitwidths}
+experiment, there are no fluctuations at all because for the lower precisions,
+there was no new model trained.
 \begin{figure}[H]
 	\centering
-	\includegraphics[width=0.75\textwidth]{figures/inferenceOnly.png}
-	\caption{Inference accuracy using float64-trained weights followed by quantization.}
+	\includegraphics[width=0.85\textwidth]{figures/inferenceOnly.png}
+	\caption{Inference accuracy using float64-trained weights followed by quantization.}\label{fig:inferenceOnlyTesting}
 \end{figure}
@@
 \subsection*{Automatic Mixed Precision and Gradient Clipping}
+As a next experiment, \dots %Sayed, also use the other images?
 \begin{figure}[H]
 	\centering
-	\includegraphics[width=0.75\textwidth]{figures/gradclip_32.png}
+	\includegraphics[width=0.85\textwidth]{figures/gradclip_32.png}
 	\caption{Inference accuracy using float32-trained weights with AMP and gradient clipping.}
@@
 \subsection*{Effect of Decimal Rounding}
-
 \begin{figure}[H]
 	\centering
-	\includegraphics[width=0.75\textwidth]{figures/118_accuracy.png}
+	\includegraphics[width=0.85\textwidth]{figures/118_accuracy.png}
 	\caption{Accuracy vs.\ number of decimal digits retained in weights.}
@@
 
+\section{Summary}
+\begin{itemize}
+	\item \textbf{16-bit is enough for this classification task}: End-to-end float16 preserved the accuracy.
+	\item \textbf{Three decimals suffice}: Storing weights down to fractional length $3$ caused no visible accuracy loss.
+\end{itemize}
+
 \section{Discussion}
 
-Our results indicate that networks trained and evaluated with as low as 16-bit
-precision can achieve comparable performance to their high-precision
-counterparts. Moreover, rounding pretrained weights to just 3 decimal digits
-preserves full accuracy, opening the door to aggressive post-training
-quantization.
+The results of the first experiments have shown that either the selected task
+(MNIST classification) was not complex enough to be affected by the decrease in
+accuracy, or the accuracy would need to be decreased even further than float16.
+
+Otherwise, the results generally indicate that networks trained and evaluated
+with as low as 16-bit precision can achieve comparable performance to their
+high-precision counterparts. Moreover, rounding pretrained weights to just 3
+decimal digits preserves full accuracy, opening the door to aggressive
+post-training quantization.
 
 This is particularly relevant for edge devices and real-time applications where
-memory and compute constraints are paramount.
+memory and compute constraints are more relevant.
 

### 2025-07-16 paper/main.tex
added decimal roudning content

@@
 \subsection*{Effect of Decimal Rounding}
+
+A non training related issue is the storage of the NN parameters. Usually, the parameters
+are stored in the format that was chosen to train with. This is best for reproducibility
+as well as accuracy when doing inference with the trained model weights. The datatype  
+precision has a higher impact during training though, which begs the question if one could
+omit precision from the model weights. In this section, we explore a naive quantization
+method: Rounding to a fixed number of decimals. With significantly little decimals and a
+reasonable range for the integer part, one could store only the significant and provide
+base as well as exponent as metadata.
+
+The first step is to examine the distribution of parameter values. By the standard
+deviation around modes one can make an educated guess as to how many decimals are needed.
+The range will decide how large the integer part may get. Reading from the 
+distribution is just a heuristic though, so we will also test the hypotheses.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.85\textwidth]{figures/118_dist.png}
+	\caption{
+		Distribution of model weights in each layer. The values are in $[-0.5, 0.5]$,
+		this means no integer part is needed in this example. Generally, with normalized data
+		and the quite popular L1-/L2-/Batch-Norm layers, weight values in $[-1, 1]$ is to be expected.}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.85\textwidth]{figures/118_dist2.png}
+	\caption{Distribution of model weights in each layer after rounding to two decimals.
+	This appears to be a good approximation already.}
+	\label{fig:118_2}
+\end{figure}
+
 \begin{figure}[H]
@@
 	\includegraphics[width=0.85\textwidth]{figures/118_accuracy.png}
-	\caption{Accuracy vs.\ number of decimal digits retained in weights.}
+	\caption{Test Accuracy vs.\ number of decimal digits retained in weights.}
+	\label{fig:118_acc}
 \end{figure}
+
+Figure~\ref{fig:118_acc} shows how rounding the decimal part affects the test accuracy.
+For just one decimal the model has a maximum of 19 different weights at its disposal,
+which makes the model have only 10\% accuracy, random guessing with a 10 class problem.
+Two decimals perform significantly better. At 70\% accuracy the model gets close to the
+full precision 88\% accuracy with 32 bit floating point. Making and educated guess on
+distribution such as Figure~\ref{fig:118_2} seems to be a good heuristic for this example.
+At just 3 decimals, the full 32 bit precision is reached. Thus we can conclude that with
+just 10 bit for representing 3 decimals, no bits for the integer part and one additional
+bit for the sign, only 11 bit must be stored to have the best performing model for
+inference.
 

### 2025-07-16 paper/main.tex
added decimal roudning content
fluffed up summary and conclusion

@@
 reasonable range for the integer part, one could store only the significant and provide
-base as well as exponent as metadata.
+base as well as exponent as metadata. When loading the model, the weights are converted
+back to floating point, with whatever precision is desired. Only the weights are quantized;
+the input data, as well are the values propagated through the network are at full
+precision.
 
@@
 \section{Summary}
+% \begin{itemize}
+% 	\item \textbf{16-bit is enough for this classification task}: End-to-end float16 preserved the accuracy.
+% 	\item \textbf{Three decimals suffice}: Storing weights down to fractional length $3$ caused no visible accuracy loss.
+% \end{itemize}
+
+% \section{Discussion}
+% The results of the first experiments have shown that either the selected task (MNIST
+% classification) was not complex enough to be affected by the decrease in accuracy, or the
+% accuracy would need to be decreased even further than float16. Otherwise, the results
+% generally indicate that networks trained and evaluated with as low as 16-bit precision can
+% achieve comparable performance to their high-precision counterparts. Moreover, rounding
+% pretrained weights to just 3 decimal digits preserves full accuracy, opening the door to
+% aggressive post-training quantization.This is particularly relevant for edge devices and
+% real-time applications where memory and compute constraints are more relevant.
+
+The experimental results presented in this paper demonstrate the robustness of neural
+networks to reduced parameter precision, particularly for simpler tasks like MNIST
+classification. Key findings include:
+
 \begin{itemize}
-	\item \textbf{16-bit is enough for this classification task}: End-to-end float16 preserved the accuracy.
-	\item \textbf{Three decimals suffice}: Storing weights down to fractional length $3$ caused no visible accuracy loss.
+	\item \textbf{Precision Reduction During Training and Inference}: Training and evaluating
+	models with 16-bit floating-point (FP16) precision preserved accuracy comparable to
+	higher-precision formats (FP32 and FP64). This suggests that for tasks of similar
+	complexity, FP16 is sufficient, offering potential memory and computational savings
+	without sacrificing performance.
+
+	\item \textbf{Post-Training Quantization}: Reducing precision only during inference, after
+	training with high-precision (FP64), maintained baseline accuracy. This is
+	particularly valuable for deployment scenarios where training is done on powerful
+	hardware, but inference occurs on resource-constrained devices.
+
+	\item \textbf{Mixed-Precision:} Training and Gradient Clipping: TODO
+
+	\item \textbf{Weight Rounding:} Rounding trained weights to just three decimal digits
+	(approximately 11 bits) achieved full accuracy, highlighting the potential for
+	aggressive quantization in storage-efficient model deployment. This finding is
+	significant for edge applications where memory footprint is critical.
 \end{itemize}
 
-\section{Discussion}
+However, the simplicity of the MNIST dataset and the small network architecture used in
+the experiments may limit the generalizability of these findings. More complex tasks
+(e.g., CIFAR-10, ImageNet) and deeper architectures (e.g., CNNs, Transformers) could
+exhibit greater sensitivity to precision reduction. Additionally, the impact of precision
+on training dynamics, such as gradient flow and convergence, was not deeply explored and
+could be a focus for future work.
 
-The results of the first experiments have shown that either the selected task
-(MNIST classification) was not complex enough to be affected by the decrease in
-accuracy, or the accuracy would need to be decreased even further than float16.
+\section{Conclusion}
 
-Otherwise, the results generally indicate that networks trained and evaluated
-with as low as 16-bit precision can achieve comparable performance to their
-high-precision counterparts. Moreover, rounding pretrained weights to just 3
-decimal digits preserves full accuracy, opening the door to aggressive
-post-training quantization.
+% This work confirms that for relatively simple tasks like MNIST classification,
+% significant reductions in parameter precision can be achieved without
+% compromising performance. Our future efforts will extend this analysis to more
+% complex datasets (e.g., CIFAR-10, ImageNet) and deeper architectures, and
+% explore advanced quantization-aware training techniques.
 
-This is particularly relevant for edge devices and real-time applications where
-memory and compute constraints are more relevant.
+This study demonstrates that neural networks can maintain strong performance even under
+significant reductions in parameter precision, particularly for simpler tasks like MNIST
+classification. Key findings confirm that FP16 training and inference, post-training
+quantization, and aggressive weight rounding (down to three decimal digits) preserve model
+accuracy while improving memory and computational efficiency. These results are promising
+for deploying models in resource-constrained environments, such as edge devices and
+real-time applications.
 
-\section{Conclusion and Future Directions}
+To build on these insights and expand their applicability, future work should focus on the
+following directions:
 
-This work confirms that for relatively simple tasks like MNIST classification,
-significant reductions in parameter precision can be achieved without
-compromising performance. Our future efforts will extend this analysis to more
-complex datasets (e.g., CIFAR-10, ImageNet) and deeper architectures, and
-explore advanced quantization-aware training techniques.
+\begin{itemize}
+\item \textbf{Complex Architectures and Datasets:} Future research should validate these
+findings on more complex models (e.g., CNNs, Transformers) and datasets (e.g., CIFAR-10,
+ImageNet) to assess the scalability of low-precision techniques.
+
+\item \textbf{Quantization-Aware Training:} Investigating advanced methods like learned
+quantization, non-uniform bit allocation, and dynamic precision adjustment could further
+optimize the trade-off between efficiency and accuracy.
+
+\item \textbf{Hardware-Specific Optimizations:} Exploring how reduced precision interacts
+with specialized hardware (e.g., TPUs, neuromorphic chips) could unlock new efficiency
+gains in real-world deployments.
+
+\item \textbf{Theoretical Understanding:} A deeper analysis of how precision affects
+gradient dynamics, convergence, and generalization could provide foundational insights for
+designing robust low-precision training algorithms.
+\end{itemize}
 

### 2025-07-16 paper/main.tex
Update main.tex

@@
 \subsection*{Automatic Mixed Precision and Gradient Clipping}
-As a next experiment, \dots %Sayed, also use the other images?
+Automatic Mixed Precision (AMP) accelerates deep learning training by using both float16
+and float32 data types reducing memory usage and increasing computational speed. 
+However, float16’s limited dynamic range can cause numerical instability,
+especially during backpropagation. To address this, Gradient Clipping is used to cap 
+the magnitude of gradients—either globally or per element—preventing extreme updates
+that could hinder convergence. Together, AMP and gradient clipping enable efficient,
+stable training by balancing performance and numerical precision.
+
+First step is using global norm clipping, all gradients are collectively scaled to ensure
+their combined norm does not exceed a specified threshold, promoting stable updates
+across the network. In second step per-element clipping constrains each individual gradient
+component within a fixed range, directly limiting extreme values.
+These methods are evaluated under Automatic Mixed Precision (AMP) to compare their impact on
+training stability, model accuracy, and computation time across different precision types.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.85\textwidth]{figures/FP16_32_clipping.png}
+	\caption{Affect of Gradient Norm on accuracy, training and testing timings.
+                 Accuracy is virtually the same between FP16 and FP32 when gradient clipping is used
+		 validating the stability of AMP and gradient scaling.
+                 Eval time is nearly identical for both formats, as evaluation involves fewer operations
+		 and less intense memory bandwidth usage.}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.85\textwidth]{figures/gradclip_16.png}
+	\caption{Effect of Per-Element Gradient Clipping on FP16.
+                 Accuracy vs Training & Eval Time}
+        \label{fig:grad16}
+\end{figure}
+
 \begin{figure}[H]
@@
 	\includegraphics[width=0.85\textwidth]{figures/gradclip_32.png}
-	\caption{Inference accuracy using float32-trained weights with AMP and gradient clipping.}
+	\caption{Effect of Per-Element Gradient Clipping on FP32.
+                 Accuracy vs Training & Eval Time}
+        \label{fig:grad32}
 \end{figure}
+
+Figure~\ref{fig:grad16} shows accuracy peaked at 97.07% with a clipping value of 0.5, 
+while the lowest accuracy (96.55%) occurred at clip=0.01, suggesting that very small
+clip values may overly restrict gradient updates and hurt learning.
+Training time ranged indicating that moderate clipping supports efficient convergence.
+Evaluation time remained stable across all clip values (around 2.01s to 2.16s),
+showing that clipping primarily impacts training rather than inference performance. 
+Also Figure~\ref{fig:grad32} shows that accuracy was most stable across all clip values
+with the highest accuracy at clip=0.5, demonstrating its robustness to gradient clipping
+variations. Training time suggesting that larger clip values can speed up convergence
+without sacrificing stability.
+Evaluation time confirming that clipping had minimal effect on inference time.
 

### 2025-07-16 paper/main.tex
fix

@@
 \subsection*{Automatic Mixed Precision and Gradient Clipping}
-Automatic Mixed Precision (AMP) accelerates deep learning training by using both float16
-and float32 data types reducing memory usage and increasing computational speed. 
-However, float16’s limited dynamic range can cause numerical instability,
-especially during backpropagation. To address this, Gradient Clipping is used to cap 
-the magnitude of gradients—either globally or per element—preventing extreme updates
-that could hinder convergence. Together, AMP and gradient clipping enable efficient,
-stable training by balancing performance and numerical precision.
+Automatic Mixed Precision (AMP) accelerates deep learning training by using
+both float16 and float32 data types reducing memory usage and increasing
+computational speed. However, float16’s limited dynamic range can cause
+numerical instability, especially during backpropagation. To address this,
+Gradient Clipping is used to cap the magnitude of gradients—either globally or
+per element—preventing extreme updates that could hinder convergence. Together,
+AMP and gradient clipping enable efficient, stable training by balancing
+performance and numerical precision.
 
-First step is using global norm clipping, all gradients are collectively scaled to ensure
-their combined norm does not exceed a specified threshold, promoting stable updates
-across the network. In second step per-element clipping constrains each individual gradient
-component within a fixed range, directly limiting extreme values.
-These methods are evaluated under Automatic Mixed Precision (AMP) to compare their impact on
-training stability, model accuracy, and computation time across different precision types.
+First step is using global norm clipping, all gradients are collectively scaled
+to ensure their combined norm does not exceed a specified threshold, promoting
+stable updates across the network. In second step per-element clipping
+constrains each individual gradient component within a fixed range, directly
+limiting extreme values. These methods are evaluated under Automatic Mixed
+Precision (AMP) to compare their impact on training stability, model accuracy,
+and computation time across different precision types.
 
@@
 
-Figure~\ref{fig:grad16} shows accuracy peaked at 97.07% with a clipping value of 0.5, 
-while the lowest accuracy (96.55%) occurred at clip=0.01, suggesting that very small
-clip values may overly restrict gradient updates and hurt learning.
-Training time ranged indicating that moderate clipping supports efficient convergence.
-Evaluation time remained stable across all clip values (around 2.01s to 2.16s),
-showing that clipping primarily impacts training rather than inference performance. 
-Also Figure~\ref{fig:grad32} shows that accuracy was most stable across all clip values
-with the highest accuracy at clip=0.5, demonstrating its robustness to gradient clipping
-variations. Training time suggesting that larger clip values can speed up convergence
-without sacrificing stability.
-Evaluation time confirming that clipping had minimal effect on inference time.
+Figure~\ref{fig:grad16} shows accuracy peaked at 97.07\% with a clipping value
+of 0.5, while the lowest accuracy (96.55\%) occurred at clip=0.01, suggesting
+that very small clip values may overly restrict gradient updates and hurt
+learning. Training time ranged indicating that moderate clipping supports
+efficient convergence. Evaluation time remained stable across all clip values
+(around 2.01s to 2.16s), showing that clipping primarily impacts training
+rather than inference performance. Also Figure~\ref{fig:grad32} shows that
+accuracy was most stable across all clip values with the highest accuracy at
+clip=0.5, demonstrating its robustness to gradient clipping variations.
+Training time suggesting that larger clip values can speed up convergence
+without sacrificing stability. Evaluation time confirming that clipping had
+minimal effect on inference time.
 
@@
 
-A non training related issue is the storage of the NN parameters. Usually, the parameters
-are stored in the format that was chosen to train with. This is best for reproducibility
-as well as accuracy when doing inference with the trained model weights. The datatype  
-precision has a higher impact during training though, which begs the question if one could
-omit precision from the model weights. In this section, we explore a naive quantization
-method: Rounding to a fixed number of decimals. With significantly little decimals and a
-reasonable range for the integer part, one could store only the significant and provide
-base as well as exponent as metadata. When loading the model, the weights are converted
-back to floating point, with whatever precision is desired. Only the weights are quantized;
-the input data, as well are the values propagated through the network are at full
-precision.
+A non training related issue is the storage of the NN parameters. Usually, the
+parameters are stored in the format that was chosen to train with. This is best
+for reproducibility as well as accuracy when doing inference with the trained
+model weights. The datatype precision has a higher impact during training
+though, which begs the question if one could omit precision from the model
+weights. In this section, we explore a naive quantization method: Rounding to a
+fixed number of decimals. With significantly little decimals and a reasonable
+range for the integer part, one could store only the significant and provide
+base as well as exponent as metadata. When loading the model, the weights are
+converted back to floating point, with whatever precision is desired. Only the
+weights are quantized; the input data, as well are the values propagated
+through the network are at full precision.
 
-The first step is to examine the distribution of parameter values. By the standard
-deviation around modes one can make an educated guess as to how many decimals are needed.
-The range will decide how large the integer part may get. Reading from the 
-distribution is just a heuristic though, so we will also test the hypotheses.
+The first step is to examine the distribution of parameter values. By the
+standard deviation around modes one can make an educated guess as to how many
+decimals are needed. The range will decide how large the integer part may get.
+Reading from the distribution is just a heuristic though, so we will also test
+the hypotheses.
 
@@
 
-Figure~\ref{fig:118_acc} shows how rounding the decimal part affects the test accuracy.
-For just one decimal the model has a maximum of 19 different weights at its disposal,
-which makes the model have only 10\% accuracy, random guessing with a 10 class problem.
-Two decimals perform significantly better. At 70\% accuracy the model gets close to the
-full precision 88\% accuracy with 32 bit floating point. Making and educated guess on
-distribution such as Figure~\ref{fig:118_2} seems to be a good heuristic for this example.
-At just 3 decimals, the full 32 bit precision is reached. Thus we can conclude that with
-just 10 bit for representing 3 decimals, no bits for the integer part and one additional
-bit for the sign, only 11 bit must be stored to have the best performing model for
-inference.
+Figure~\ref{fig:118_acc} shows how rounding the decimal part affects the test
+accuracy. For just one decimal the model has a maximum of 19 different weights
+at its disposal, which makes the model have only 10\% accuracy, random guessing
+with a 10 class problem. Two decimals perform significantly better. At 70\%
+accuracy the model gets close to the full precision 88\% accuracy with 32 bit
+floating point. Making and educated guess on distribution such as
+Figure~\ref{fig:118_2} seems to be a good heuristic for this example. At just 3
+decimals, the full 32 bit precision is reached. Thus we can conclude that with
+just 10 bit for representing 3 decimals, no bits for the integer part and one
+additional bit for the sign, only 11 bit must be stored to have the best
+performing model for inference.
 
@@
 
-The experimental results presented in this paper demonstrate the robustness of neural
-networks to reduced parameter precision, particularly for simpler tasks like MNIST
-classification. Key findings include:
+The experimental results presented in this paper demonstrate the robustness of
+neural networks to reduced parameter precision, particularly for simpler tasks
+like MNIST classification. Key findings include:
 
@@
 
-However, the simplicity of the MNIST dataset and the small network architecture used in
-the experiments may limit the generalizability of these findings. More complex tasks
-(e.g., CIFAR-10, ImageNet) and deeper architectures (e.g., CNNs, Transformers) could
-exhibit greater sensitivity to precision reduction. Additionally, the impact of precision
-on training dynamics, such as gradient flow and convergence, was not deeply explored and
-could be a focus for future work.
+However, the simplicity of the MNIST dataset and the small network architecture
+used in the experiments may limit the generalizability of these findings. More
+complex tasks (e.g., CIFAR-10, ImageNet) and deeper architectures (e.g., CNNs,
+Transformers) could exhibit greater sensitivity to precision reduction.
+Additionally, the impact of precision on training dynamics, such as gradient
+flow and convergence, was not deeply explored and could be a focus for future
+work.
 
@@
 
-This study demonstrates that neural networks can maintain strong performance even under
-significant reductions in parameter precision, particularly for simpler tasks like MNIST
-classification. Key findings confirm that FP16 training and inference, post-training
-quantization, and aggressive weight rounding (down to three decimal digits) preserve model
-accuracy while improving memory and computational efficiency. These results are promising
-for deploying models in resource-constrained environments, such as edge devices and
-real-time applications.
+This study demonstrates that neural networks can maintain strong performance
+even under significant reductions in parameter precision, particularly for
+simpler tasks like MNIST classification. Key findings confirm that FP16
+training and inference, post-training quantization, and aggressive weight
+rounding (down to three decimal digits) preserve model accuracy while improving
+memory and computational efficiency. These results are promising for deploying
+models in resource-constrained environments, such as edge devices and real-time
+applications.
 
-To build on these insights and expand their applicability, future work should focus on the
-following directions:
+To build on these insights and expand their applicability, future work should
+focus on the following directions:
 

### 2025-07-16 paper/main.tex
almost done

@@
 	\centering
-	\includegraphics[width=0.85\textwidth]{figures/FP16_32_clipping.png}
+	\includegraphics[width=0.95\textwidth]{figures/FP16_32_clipping.png}
 	\caption{Affect of Gradient Norm on accuracy, training and testing timings.
@@
 	\centering
-	\includegraphics[width=0.85\textwidth]{figures/gradclip_16.png}
-	\caption{Effect of Per-Element Gradient Clipping on FP16.
-		Accuracy vs Training & Eval Time}
-	\label{fig:grad16}
+	\includegraphics[width=0.95\textwidth]{figures/gradclip_16.png}
+	\caption{Effect of Per-Element Gradient Clipping on FP16. Accuracy vs Training \& Eval Time}\label{fig:grad16}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.85\textwidth]{figures/gradclip_32.png}
-	\caption{Effect of Per-Element Gradient Clipping on FP32.
-		Accuracy vs Training & Eval Time}
-	\label{fig:grad32}
+	\includegraphics[width=0.95\textwidth]{figures/gradclip_32.png}
+	\caption{Effect of Per-Element Gradient Clipping on FP32. Accuracy vs Training \& Eval Time}\label{fig:grad32}
 \end{figure}
@@
 	\caption{Distribution of model weights in each layer after rounding to two decimals.
-		This appears to be a good approximation already.}
-	\label{fig:118_2}
+		This appears to be a good approximation already.}\label{fig:118_2}
 \end{figure}
@@
 	\includegraphics[width=0.85\textwidth]{figures/118_accuracy.png}
-	\caption{Test Accuracy vs.\ number of decimal digits retained in weights.}
-	\label{fig:118_acc}
+	\caption{Test Accuracy vs.\ number of decimal digits retained in weights.}\label{fig:118_acc}
 \end{figure}

### 2025-07-16 paper/main.tex
almost done

@@
 	We study how the numerical precision of weights, activations, and gradients in neural networks can be reduced without sacrificing too much of the classification performance.
-Using MNIST as a dataset, we benchmark uniform 16–64-bit quantisation, mixed-precision formats, and post-training decimal rounding.
+Using MNIST as a dataset, we benchmark uniform 64-16-bit quantisation, mixed-precision formats, and post-training decimal rounding.
 The results reveal that (i) 16-bit end-to-end training preserves full accuracy for the MNIST dataset and the network used, (ii) three-decimal-place weight rounding while keeping 32-bit precision incurs minimal accuracy loss.
@@
 numerical instability, especially during backpropagation. To address this,
-Gradient Clipping is used to cap the magnitude of gradients—either globally or
-per element—preventing extreme updates that could hinder convergence. Together,
-AMP and gradient clipping enable efficient, stable training by balancing
-performance and numerical precision.
+Gradient Clipping is used to cap the magnitude of gradients (either globally or
+per element) preventing extreme updates that could hinder convergence.
+Together, AMP and gradient clipping enable efficient, stable training by
+balancing performance and numerical precision.
 

### 2025-07-16 paper/main.tex
completed summary

@@
 \begin{itemize}
-	\item \textbf{Precision Reduction During Training and Inference}: Training and evaluating
-	      models with 16-bit floating-point (FP16) precision preserved accuracy comparable to
-	      higher-precision formats (FP32 and FP64). This suggests that for tasks of similar
-	      complexity, FP16 is sufficient, offering potential memory and computational savings
-	      without sacrificing performance.
+	\item \textbf{Precision Reduction During Training and Inference}: Training and
+	      evaluating models with 16-bit floating-point (FP16) precision preserved accuracy
+	      comparable to higher-precision formats (FP32 and FP64). This suggests that for
+	      tasks of similar complexity, FP16 is sufficient, offering potential memory and
+	      computational savings without sacrificing performance.
 
@@
 
-	\item \textbf{Mixed-Precision:} Training and Gradient Clipping: TODO
+	\item \textbf{Mixed-Precision:} Training with mixed precision and Gradient Clipping
+		  resulted in similar test accuracy scores as precision reduction during training
+		  as well as post-training quantization. No significant decrease in model
+		  performance accuracy was detected. The train and test times were similar,
+		  regardless of using FP32, FP16 or a mix of both.
 

### 2025-07-16 paper/main.tex
layout

@@
 \begin{abstract}
-	We study how the numerical precision of weights, activations, and gradients in neural networks can be reduced without sacrificing too much of the classification performance.
-Using MNIST as a dataset, we benchmark uniform 64-16-bit quantisation, mixed-precision formats, and post-training decimal rounding.
-The results reveal that (i) 16-bit end-to-end training preserves full accuracy for the MNIST dataset and the network used, (ii) three-decimal-place weight rounding while keeping 32-bit precision incurs minimal accuracy loss.
-These findings verify limits for memory- and energy-constrained usecases like inference in mobile devices.
+	We study how the numerical precision of weights, activations, and gradients in neural
+	networks can be reduced without sacrificing too much of the classification
+	performance. Using MNIST as a dataset, we benchmark uniform 64-16-bit quantisation,
+	mixed-precision formats, and post-training decimal rounding. The results reveal that
+	(i) 16-bit end-to-end training does not significantly decrease test accuracy for the
+	MNIST dataset and the network used, (ii) three-decimal-place weight rounding while
+	keeping 32-bit precision incurs minimal accuracy loss. These findings verify limits
+	for memory- and energy-constrained usecases like inference in mobile devices.
 \end{abstract}
@@
 
+\newpage
+
 \section{Experimental Design}
@@
 
+\newpage
 \section{Results}

### 2025-07-16 paper/main.tex
fertig

@@
 \begin{itemize}
-	\item \textbf{Mixed Precision Training}, which leverages float16 arithmetic for speed while maintaining float32 master weights~\cite{micikevicius2017mixed}.
+	\item \textbf{Mixed Precision Training}, which leverages FP16 arithmetic for speed while maintaining FP32 master weights~\cite{micikevicius2017mixed}.
 	\item \textbf{DoReFa-Net}, which quantizes weights, activations, and gradients to arbitrary bitwidths~\cite{zhou2018dorefa}.
@@
 includes the weights and biases, activations, gradients, optimizer states, etc.
-The types tried were float64, float32, and float16.
+The types tried were FP64, FP32, and FP16.
 
@@
 y-axis is not starting from zero, so the absolute height differences are
-amplified visually, though the delta is only around $.1$ percent. Re-training
-the models and re-evaluating them has also lead to the accuracies being
-``reversed'', meaning the float64 model had the least accuracy. It is safe to
+amplified visually, though the delta is only around $.1\%$. Re-training the
+models and re-evaluating them has also lead to the accuracies being
+``reversed'', meaning the FP64 model had the least accuracy. It is safe to
 assume that the model accuracy has neither decreased, nor increased.
@@
 This third experiment was conducted to test a possible post-training precision
-reduction. For this, the model was trained on the highest precision (float64)
-and then quantized at the inference stage.
-Figure~\ref{fig:inferenceOnlyTesting} shows that the accuracy of the baseline
-(float64 training AND inference) is maintained as the precision is decreased at
-the later stages. In contrast to the \textit{Accuracy Across Bitwidths}
-experiment, there are no fluctuations at all because for the lower precisions,
-there was no new model trained.
+reduction. For this, the model was trained on the highest precision (FP64) and
+then quantized at the inference stage. Figure~\ref{fig:inferenceOnlyTesting}
+shows that the accuracy of the baseline (FP64 training AND inference) is
+maintained as the precision is decreased at the later stages. In contrast to
+the \textit{Accuracy Across Bitwidths} experiment, there are no fluctuations at
+all because for the lower precisions, there was no new model trained.
 \begin{figure}[H]
@@
 	\includegraphics[width=0.85\textwidth]{figures/inferenceOnly.png}
-	\caption{Inference accuracy using float64-trained weights followed by quantization.}\label{fig:inferenceOnlyTesting}
+	\caption{Inference accuracy using FP64-trained weights followed by quantization.}\label{fig:inferenceOnlyTesting}
 \end{figure}
@@
 Automatic Mixed Precision (AMP) accelerates deep learning training by using
-both float16 and float32 data types reducing memory usage and increasing
-computational speed. However, float16’s limited dynamic range can cause
-numerical instability, especially during backpropagation. To address this,
-Gradient Clipping is used to cap the magnitude of gradients (either globally or
-per element) preventing extreme updates that could hinder convergence.
-Together, AMP and gradient clipping enable efficient, stable training by
-balancing performance and numerical precision.
+both FP16 and FP32 data types reducing memory usage and increasing
+computational speed. However, FP16’s limited dynamic range can cause numerical
+instability, especially during backpropagation. To address this, Gradient
+Clipping is used to cap the magnitude of gradients (either globally or per
+element) preventing extreme updates that could hinder convergence. Together,
+AMP and gradient clipping enable efficient, stable training by balancing
+performance and numerical precision.
 
@@
 % \begin{itemize}
-% 	\item \textbf{16-bit is enough for this classification task}: End-to-end float16 preserved the accuracy.
+% 	\item \textbf{16-bit is enough for this classification task}: End-to-end FP16 preserved the accuracy.
 % 	\item \textbf{Three decimals suffice}: Storing weights down to fractional length $3$ caused no visible accuracy loss.
@@
 % classification) was not complex enough to be affected by the decrease in accuracy, or the
-% accuracy would need to be decreased even further than float16. Otherwise, the results
+% accuracy would need to be decreased even further than FP16. Otherwise, the results
 % generally indicate that networks trained and evaluated with as low as 16-bit precision can

### 2025-11-04 paper/main.tex
weiter

@@
 
-% Geometry settings
 \geometry{
@@
 
-% Title settings
 \titleformat{\section}{\normalfont\Large\bfseries}{\thesection}{1em}{}
@@
 
-% Title
 \title{An Experimental Analysis of the Influence of the Precision of Parameters in an Artificial Neural Network}
@@
 \begin{abstract}
-	We study how the numerical precision of weights, activations, and gradients in neural
-	networks can be reduced without sacrificing too much of the classification
-	performance. Using MNIST as a dataset, we benchmark uniform 64-16-bit quantisation,
-	mixed-precision formats, and post-training decimal rounding. The results reveal that
-	(i) 16-bit end-to-end training does not significantly decrease test accuracy for the
-	MNIST dataset and the network used, (ii) three-decimal-place weight rounding while
-	keeping 32-bit precision incurs minimal accuracy loss. These findings verify limits
-	for memory- and energy-constrained usecases like inference in mobile devices.
+	\noindent
+	We study how the numerical precision of weights, activations, and gradients in
+	neural networks can be reduced without sacrificing too much of the performance.
+	Using MNIST as a dataset, we benchmark uniform quantisation (mapping real
+	values to evenly spaced levels via a single step size), mixed-precision
+	formats, and post-training decimal rounding. We compare end-to-end
+	low-precision training, inference-only quantisation, and mixed-precision with
+	gradient clipping, and evaluate their impact on test accuracy and efficiency.
+	The results on the used test setup show that (i) 16-bit end-to-end training
+	does not significantly decrease test accuracy for the MNIST dataset, (ii)
+	three-decimal-place weight rounding while keeping 32-bit precision incurs
+	minimal accuracy loss. These findings verify limits for memory- and
+	energy-constrained usecases like inference in mobile devices.
 \end{abstract}
@@
 \section{Introduction}
-Artificial Neural Networks (ANNs) are a key part of today’s machine learning
-systems. They’ve made huge progress in areas like image recognition, speech
+Artificial Neural Networks (ANNs) are a key part of today's machine learning
+systems. They have made huge progress in areas like image recognition, speech
 processing, and self-driving cars. However, one of the biggest challenges with
-ANNs is that they use a lot of memory and computing power. This becomes a
+large ANNs is that they use a lot of memory and computing power. This becomes a
 problem when you want to run them on devices with limited resources, such as
-smartphones, embedded systems, or edge devices. \\ \\ To solve this,
+smartphones, embedded systems, or edge devices. \\\\ In order to solve this,
 researchers have looked into ways to make neural networks more efficient. A
 common approach is to reduce the size of the network or shrink the input batch
-size. But doing this can harm the network’s ability to learn and generalize. A
-better strategy is to reduce the numerical precision of the network’s
-parameters such as things like weights, activations, and gradients. This
-method, called quantization, makes the model smaller and faster while using
-less energy, without necessarily hurting accuracy. \\ \\ One advanced version
-of this method is called mixed-precision quantization. Instead of using the
-same number of bits for every layer in the network, mixed-precision assigns
-different bitwidths depending on how important each part is. For example, the
-early layers that handle low-level features might use higher precision, while
-deeper layers could work fine with fewer bits. This flexible approach often
-leads to better results than using a single fixed precision for the whole
-model. Researchers have used a range of techniques, like gradient descent,
-heuristics, and even evolutionary algorithms, to find the best bitwidth
-settings for each layer. \\ \\ Another method that helps reduce resource usage
-is network pruning. This means cutting out parts of the network that don’t
-contribute much to its performance. One interesting way to do this is with
-simulated annealing, a search algorithm that mimics the process of cooling
-metal to find a stable structure. It doesn’t need gradient information and can
-still find good network shapes by exploring many possibilities. This makes it
-useful when you want to shrink a network without going through heavy
-retraining. \\ \\ Previous studies have shown that reducing the precision of
-neural network parameters can lead to significant improvements in memory usage
-and processing speed. Even when precision is reduced to very low levels, models
-can still perform competitively on complex tasks. However, while these early
-results are promising, the overall effects of lowering precision, especially
-how it influences model accuracy, training behavior, and storage efficiency,
-are still not fully understood and deserve more detailed investigation. \\ \\
-This is where our experimental work begins. In this paper, we take a closer
-look at how different levels of parameter precision affect the behavior and
-performance of neural networks. Our goal is to understand what happens when we
-reduce precision in various ways, and how much we can lower it before
-performance starts to drop. \\ \\ We start by testing how reducing precision
-during both training and inference affects model accuracy and runtime. Next, we
-look at what happens when we train the network using full precision, but lower
-the precision only during inference. This setup is common in real-world
-applications, where a powerful machine handles training, but the final model is
-deployed on a lightweight device. Then, we run a series of tests where we round
-the trained weights to a limited number of decimal places. This helps us see
-how much we can simplify the stored weight values without hurting accuracy too
-much. This is especially useful for devices with very limited storage. We also
-try out mixed-precision training, using a mix of FP16 and FP32 formats. On top
-of that, we experiment with different types of gradient clipping, like global
-norm clipping and value clipping, to see how they affect training stability and
-final accuracy. Even though we use relatively simple network architectures and
-datasets, our results show that neural networks are quite robust to
-low-precision settings. In some cases, we were able to round weights to just
-three decimal digits and still get nearly the same accuracy as the
-full-precision version. These findings show that it’s possible to significantly
-reduce the memory and compute needs of neural networks without losing much
-performance. This makes them more practical for use in low-resource settings.
-It also opens the door for more advanced techniques that combine quantization,
-pruning, and precision-aware training to build highly efficient models for
-real-world applications.
+size. But doing this can harm the network's ability to learn and generalize. A
+better strategy is to reduce the numerical precision of the network's
+parameters such as weights, activations, and gradients. This method, called
+quantization, makes the model smaller and faster while using less energy,
+without necessarily hurting accuracy. \\\\ An advanced version of this method
+is called mixed-precision quantization. Instead of using the same number of
+bits for every layer in the network, mixed-precision assigns different
+bitwidths depending on how important each part is. For example, the early
+layers that handle low-level features might use higher precision, while deeper
+layers could work fine with fewer bits. This flexible approach often leads to
+better results than using a single fixed precision for the whole model.
+Researchers have used a range of techniques, like gradient descent, heuristics,
+and even evolutionary algorithms, to find the best bitwidth settings for each
+layer. \\\\ Another method that helps reduce resource usage is network pruning.
+This means cutting out parts of the network that do not contribute much to its
+performance. A way to do this is simulated annealing, a search algorithm that
+mimics the process of cooling metal to find a stable structure. It does not
+need gradient information and can still find good network shapes by exploring
+many possibilities. This makes it useful when you want to shrink a network
+without going through heavy retraining. \\\\ Previous studies have shown that
+reducing the precision of neural network parameters can lead to significant
+improvements in memory usage and processing speed. Even when precision is
+reduced to very low levels, models can still perform competitively on complex
+tasks. However, while these early results are promising, the overall effects of
+lowering precision, especially how it influences model accuracy, training
+behavior, and storage efficiency, are still not fully understood and deserve
+more detailed investigation. \\\\ This is where our experimental work begins.
+In this paper, we take a closer look at how different levels of parameter
+precision affect the behavior and performance of neural networks. Our goal is
+to understand what happens when we reduce precision in various ways, and how
+much we can lower it before performance starts to drop. \\\\ We start by
+testing how reducing precision during both training and inference affects model
+accuracy and runtime. Next, we look at what happens when we train the network
+using full precision, but lower the precision only during inference. This setup
+is common in real-world applications, where a powerful machine handles
+training, but the final model is deployed on a lightweight device. Then, we run
+a series of tests where we round the trained weights to a limited number of
+decimal places. This helps to see how much we can simplify the stored weight
+values without hurting accuracy too much. This is especially useful for devices
+with very limited storage. We also try out mixed-precision training, using a
+mix of Half-Precision Floating Point (FP16) and Single-Precision Floating Point
+(FP32) formats. On top of that, we experiment with different types of gradient
+clipping, like global norm clipping and value clipping, to see how they affect
+training stability and final accuracy. \\\\ Even though we use relatively
+simple network architectures and datasets, our results show that neural
+networks are quite robust to low-precision settings. In some cases, we were
+able to round weights to just three decimal digits and still get nearly the
+same accuracy as the full-precision version. These findings show that it is
+possible to significantly reduce the memory and compute needs of neural
+networks without losing much performance. This makes them more practical for
+use in low-resource settings. It also opens the door for more advanced
+techniques that combine quantization, pruning, and precision-aware training to
+build highly efficient models for real-world applications.
 
-\section{Background}
+\section{Artificial Neural Networks}
 
 A typical neuron in a neural network performs a weighted sum of its inputs,
-followed by an activation function:
-\[
+followed by an activation function (depicted in~\eqref{eq:activation}):
+
+\begin{equation}\label{eq:activation}
 	a = \omega^\top x + b,\quad \hat{y} = \sigma(a),
-\]
-where \( \omega \in \mathbb{R}^n \) is the weight vector, \( b \in \mathbb{R}
-\) is the bias, \( x \in \mathbb{R}^n \) is the input, and \( \sigma \) is a
-nonlinear activation function (e.g., sigmoid or ReLU).
+\end{equation}
 
-Training is conducted via backpropagation, where gradients of a loss function
-\( L(\hat{y}, y) \) with respect to the model parameters are computed. For
-example, using the mean squared error:
-% \[
-% L(\hat{y}, y) = (\hat{y} - y)^2, \qquad \pdv{L}{\omega_i} = 2(\hat{y} - y) \cdot \odv{\sigma}{a} \cdot x_i.
-% \]
+where $\omega \in \mathbb{R}^n$ is the weight vector, $b \in \mathbb{R}$ is the
+bias, $x \in \mathbb{R}^n$ is the input, and $\sigma$ is a nonlinear activation
+function (e.g., sigmoid or ReLU). Training is conducted via backpropagation,
+where gradients of a loss function $L(\hat{y}, y)$ with respect to the model
+parameters are computed. For example, using the mean squared error:
 
-\begin{align*}
-	L(f(x_1, x_2)) & = {(\sigma(\omega^T x + b) - y)}^2 \\
-	               & = {(\sigma(a) - y)}^2              \\
+\begin{align}\label{eq:mse}
+	L(f(x_1, x_2)) & = {(\sigma(\omega^\top x + b) - y)}^2 \nonumber \\
+	               & = {(\sigma(a) - y)}^2 \nonumber                 \\
 	               & = {(\hat y - y)}^2
-\end{align*}
-\begin{alignat*}{3}
-	\pdv{L}{\omega_1} & = \pdv{L}{\sigma} &  & \pdv{\sigma}{a}      &  & \pdv{a}{\omega_1} \\
-	                  & = 2(\hat y - y)~~ &  & \odv{}{a}\sigma(a)~~ &  & x_1
-\end{alignat*}
+\end{align}
+\begin{alignat}{3}\label{eq:grad-w1}
+	\pdv{L}{\omega_1} & = \pdv{L}{\sigma} &  & \pdv{\sigma}{a}      &  & \pdv{a}{\omega_1} \nonumber \\
+	                  & = 2(\hat y - y)~~ &  & \odv{\sigma(a)}{a}~~ &  & x_1
+\end{alignat}
 
@@
 	\includegraphics[width=0.85\textwidth]{figures/trainingAndInference.png}
-	\caption{Model accuracy evaluated after training with the different precisions.}\label{fig:modelAccTrainAndInf}
+	\caption{Model accuracy evaluated after training with different precisions.}\label{fig:modelAccTrainAndInf}
 \end{figure}
@@
 \subsection*{Training and Inference Time}
-Also analyzed were the times needed for both the training and inference.
+We also analyzed the times needed for both, training and inference.
 Figure~\ref{fig:modelEvalTrainTimes} depicts the times measured in the training
@@
 both FP16 and FP32 data types reducing memory usage and increasing
-computational speed. However, FP16’s limited dynamic range can cause numerical
+computational speed. However, FP16's limited dynamic range can cause numerical
 instability, especially during backpropagation. To address this, Gradient

### 2025-11-09 paper/main.tex
weiter

@@
 16-bit or even 8-bit fixed-point representations.
-Figure~\ref{fig:floatingPointFormat} depicts the general floating point format
+Equation~\ref{eq:floatingPointFormat} depicts the general floating point format
 and terminology of significand, base, and exponent.
@@
 
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{figures/float.png}
-	\caption{Illustration of standard floating-point format.}\label{fig:floatingPointFormat}
-\end{figure}
+\begin{align}\label{eq:floatingPointFormat}
+	\frac{2469}{200}=12.345=\underbrace{12345}_{\text{significand}}\times\underbrace{10^{\overbrace{-3}^{\text{exponent}}}}_{\text{base}}
+\end{align}
 

### 2025-11-09 paper/main.tex
weiter

@@
 	neural networks can be reduced without sacrificing too much of the performance.
-	Using MNIST as a dataset, we benchmark uniform quantisation (mapping real
+	Using MNIST as a dataset, we benchmark uniform quantization (mapping real
 	values to evenly spaced levels via a single step size), mixed-precision
-	formats, and post-training decimal rounding. We compare end-to-end
-	low-precision training, inference-only quantisation, and mixed-precision with
+	formats, and post-training decimal rounding. We evaluate precision settings
+	(FP64/FP32/FP16) across all tensors during training, experiment with an
+	inference-only quantization path on FP64-trained models, and test automatic
+	mixed precision with gradient clipping. Metrics include test accuracy,
+	training/inference time, and parameter storage. We compare end-to-end
+	low-precision training, inference-only quantization, and mixed-precision with
 	gradient clipping, and evaluate their impact on test accuracy and efficiency.
@@
 build highly efficient models for real-world applications.
+
+\subsection{Related Work}
+
+Prior research in this domain can be broadly classified into two categories:
+
+\subsubsection*{Quantization During Training}
+Several studies have explored training neural networks with reduced precision
+directly. Notable contributions include:
+\begin{itemize}
+	\item \textbf{Mixed Precision Training}, which leverages FP16 arithmetic for speed while maintaining FP32 master weights~\cite{micikevicius2017mixed}.
+	\item \textbf{DoReFa-Net}, which quantizes weights, activations, and gradients to arbitrary bitwidths~\cite{zhou2018dorefa}.
+	\item \textbf{Q-BERT}, a quantized BERT model that maintains accuracy using only 4-bit weights~\cite{shen2020qbert}.
+\end{itemize}
+
+\subsubsection*{Post-Training Quantization}
+Other approaches focus on converting pretrained models into lower-precision
+equivalents:
+\begin{itemize}
+	\item \textbf{Integer-Only Quantization} using symmetric quantization scales~\cite{jacob2018quantization}.
+	\item \textbf{GPTQ}, a method for quantizing large transformer models without retraining~\cite{frantar2022gptq}.
+	\item \textbf{SmoothQuant}, which scales activations to better match weight ranges during quantization~\cite{xiao2024smoothquant}.
+\end{itemize}
 
@@
 \end{table}
-
-\section{Related Work}
-
-Prior research in this domain can be broadly classified into two categories:
-
-\subsection*{Quantization During Training}
-Several studies have explored training neural networks with reduced precision
-directly. Notable contributions include:
-\begin{itemize}
-	\item \textbf{Mixed Precision Training}, which leverages FP16 arithmetic for speed while maintaining FP32 master weights~\cite{micikevicius2017mixed}.
-	\item \textbf{DoReFa-Net}, which quantizes weights, activations, and gradients to arbitrary bitwidths~\cite{zhou2018dorefa}.
-	\item \textbf{Q-BERT}, a quantized BERT model that maintains accuracy using only 4-bit weights~\cite{shen2020qbert}.
-\end{itemize}
-
-\subsection*{Post-Training Quantization}
-Other approaches focus on converting pretrained models into lower-precision
-equivalents:
-\begin{itemize}
-	\item \textbf{Integer-Only Quantization} using symmetric quantization scales~\cite{jacob2018quantization}.
-	\item \textbf{GPTQ}, a method for quantizing large transformer models without retraining~\cite{frantar2022gptq}.
-	\item \textbf{SmoothQuant}, which scales activations to better match weight ranges during quantization~\cite{xiao2024smoothquant}.
-\end{itemize}
-
-\newpage
 

### 2025-11-12 paper/main.tex
weiter

@@
 	neural networks can be reduced without sacrificing too much of the performance.
-	Using MNIST as a dataset, we benchmark uniform quantization (mapping real
-	values to evenly spaced levels via a single step size), mixed-precision
-	formats, and post-training decimal rounding. We evaluate precision settings
-	(FP64/FP32/FP16) across all tensors during training, experiment with an
-	inference-only quantization path on FP64-trained models, and test automatic
-	mixed precision with gradient clipping. Metrics include test accuracy,
-	training/inference time, and parameter storage. We compare end-to-end
-	low-precision training, inference-only quantization, and mixed-precision with
-	gradient clipping, and evaluate their impact on test accuracy and efficiency.
-	The results on the used test setup show that (i) 16-bit end-to-end training
-	does not significantly decrease test accuracy for the MNIST dataset, (ii)
-	three-decimal-place weight rounding while keeping 32-bit precision incurs
-	minimal accuracy loss. These findings verify limits for memory- and
-	energy-constrained usecases like inference in mobile devices.
+	Using MNIST as a general dataset (and FashionMNIST for the decimal rounding),
+	we benchmark uniform quantization (mapping real values to evenly spaced levels
+	via a single step size), mixed-precision formats, and post-training decimal
+	rounding. We evaluate precision settings (FP64/FP32/FP16) across all tensors
+	during training, experiment with an inference-only quantization path on
+	FP64-trained models, and test automatic mixed precision with gradient clipping.
+	Metrics include test accuracy, training/inference time, and parameter storage.
+	We compare end-to-end low-precision training, inference-only quantization, and
+	mixed-precision with gradient clipping, and evaluate their impact on test
+	accuracy and efficiency. The results on the used test setup show that (i)
+	16-bit end-to-end training does not significantly decrease test accuracy for
+	the MNIST dataset, (ii) three-decimal-place weight rounding while keeping
+	32-bit precision incurs minimal accuracy loss. These findings verify limits for
+	memory- and energy-constrained usecases like inference in mobile devices.
 \end{abstract}

### 2025-11-12 paper/main.tex
weiter

@@
 better results than using a single fixed precision for the whole model.
-Researchers have used a range of techniques, like gradient descent, heuristics,
-and even evolutionary algorithms, to find the best bitwidth settings for each
-layer. \\\\ Another method that helps reduce resource usage is network pruning.
-This means cutting out parts of the network that do not contribute much to its
-performance. A way to do this is simulated annealing, a search algorithm that
-mimics the process of cooling metal to find a stable structure. It does not
-need gradient information and can still find good network shapes by exploring
-many possibilities. This makes it useful when you want to shrink a network
-without going through heavy retraining. \\\\ Previous studies have shown that
-reducing the precision of neural network parameters can lead to significant
-improvements in memory usage and processing speed. Even when precision is
-reduced to very low levels, models can still perform competitively on complex
-tasks. However, while these early results are promising, the overall effects of
-lowering precision, especially how it influences model accuracy, training
-behavior, and storage efficiency, are still not fully understood and deserve
-more detailed investigation. \\\\ This is where our experimental work begins.
-In this paper, we take a closer look at how different levels of parameter
-precision affect the behavior and performance of neural networks. Our goal is
-to understand what happens when we reduce precision in various ways, and how
-much we can lower it before performance starts to drop. \\\\ We start by
-testing how reducing precision during both training and inference affects model
-accuracy and runtime. Next, we look at what happens when we train the network
-using full precision, but lower the precision only during inference. This setup
-is common in real-world applications, where a powerful machine handles
-training, but the final model is deployed on a lightweight device. Then, we run
-a series of tests where we round the trained weights to a limited number of
-decimal places. This helps to see how much we can simplify the stored weight
-values without hurting accuracy too much. This is especially useful for devices
-with very limited storage. We also try out mixed-precision training, using a
-mix of Half-Precision Floating Point (FP16) and Single-Precision Floating Point
-(FP32) formats. On top of that, we experiment with different types of gradient
-clipping, like global norm clipping and value clipping, to see how they affect
-training stability and final accuracy. \\\\ Even though we use relatively
-simple network architectures and datasets, our results show that neural
-networks are quite robust to low-precision settings. In some cases, we were
-able to round weights to just three decimal digits and still get nearly the
-same accuracy as the full-precision version. These findings show that it is
-possible to significantly reduce the memory and compute needs of neural
-networks without losing much performance. This makes them more practical for
-use in low-resource settings. It also opens the door for more advanced
-techniques that combine quantization, pruning, and precision-aware training to
-build highly efficient models for real-world applications.
+Researchers have used a range of techniques, like stochastic gradient descent
+(SGD), heuristics, and even evolutionary algorithms, to find the best bitwidth
+settings for each layer. \\\\ Another method that helps reduce resource usage
+is network pruning. This means cutting out parts of the network that do not
+contribute much to its performance. A way to do this is simulated annealing, a
+search algorithm that mimics the process of cooling metal to find a stable
+structure. It does not need gradient information and can still find good
+network shapes by exploring many possibilities. This makes it useful when you
+want to shrink a network without going through heavy retraining. \\\\ Previous
+studies have shown that reducing the precision of neural network parameters can
+lead to significant improvements in memory usage and processing speed. Even
+when precision is reduced to very low levels, models can still perform
+competitively on complex tasks. However, while these early results are
+promising, the overall effects of lowering precision, especially how it
+influences model accuracy, training behavior, and storage efficiency, are still
+not fully understood and deserve more detailed investigation. \\\\ This is
+where our experimental work begins. In this paper, we take a closer look at how
+different levels of parameter precision affect the behavior and performance of
+neural networks. Our goal is to understand what happens when we reduce
+precision in various ways, and how much we can lower it before performance
+starts to drop. \\\\ We start by testing how reducing precision during both
+training and inference affects model accuracy and runtime. Next, we look at
+what happens when we train the network using full precision, but lower the
+precision only during inference. This setup is common in real-world
+applications, where a powerful machine handles training, but the final model is
+deployed on a lightweight device. Then, we run a series of tests where we round
+the trained weights to a limited number of decimal places. This helps to see
+how much we can simplify the stored weight values without hurting accuracy too
+much. This is especially useful for devices with very limited storage. We also
+try out mixed-precision training, using a mix of Half-Precision Floating Point
+(FP16) and Single-Precision Floating Point (FP32) formats. On top of that, we
+experiment with different types of gradient clipping, like global norm clipping
+and value clipping, to see how they affect training stability and final
+accuracy. \\\\ Even though we use relatively simple network architectures and
+datasets, our results show that neural networks are quite robust to
+low-precision settings. In some cases, we were able to round weights to just
+three decimal digits and still get nearly the same accuracy as the
+full-precision version. These findings show that it is possible to
+significantly reduce the memory and compute needs of neural networks without
+losing much performance. This makes them more practical for use in low-resource
+settings. It also opens the door for more advanced techniques that combine
+quantization, pruning, and precision-aware training to build highly efficient
+models for real-world applications.
 
@@
 
-We trained a simple fully connected neural network on the MNIST dataset,
-consisting of 60,000 training and 10,000 test grayscale images of handwritten
-digits (28×28 pixels, see figure~\ref{fig:mnistSample}). Model training was
-performed under varying parameter bitwidths, and 5-fold cross-validation to
-evaluate the model performance. The structure of the MLP is depicted in
-figure~\ref{fig:mlpStructure}.
+The effect of numerical precision on training and inference is evaluated
+through five experiment groups run on standard image classification tasks. The
+main dataset MNIST\footnote{LeCun, Cortes, and Burges:
+	\url{http://yann.lecun.com/exdb/mnist/}.} ($28\times 28$ grayscale images of
+handwritten digits, train/test split $60.000$/$10.000$, 10 classes, examples
+shown in Figure~\ref{fig:mnistSample}) is used to determine precision effects
+on accuracy and runtime under a simple workload, while an after-training
+rounding experiment is performed on FashionMNIST\footnote{Xiao, Rasul, and
+	Vollgraf: \url{https://github.com/zalandoresearch/fashion-mnist}.} ($28\times
+28$ grayscale images of clothing items, train/test split $60.000$/$10.000$, 10
+classes) so the impact of using coarse value sets is better observable.\\
 
@@
 
+The classifier used is a fully connected network
+$784\rightarrow128\rightarrow64\rightarrow10$ with ReLU between linear layers
+(Figure~\ref{fig:mlpStructure}). Inputs are normalized with mean $0.1307$ and
+std $0.3081$. Unless stated otherwise, we train with cross-entropy loss and SGD
+(learning rate $0.01$, momentum $0.9$), batch size $64$, and on CUDA.\\
+
 \begin{figure}[H]
@@
 	\includegraphics[width=0.85\textwidth]{figures/mlpStruc.png}
-	\caption{MLP structure used in the experiments.}\label{fig:mlpStructure}
+	\caption{Network structure used in the experiments.}\label{fig:mlpStructure}
 \end{figure}

### 2025-11-12 paper/main.tex
weiter

@@
 \section{Experimental Design}
-
 The effect of numerical precision on training and inference is evaluated
@@
 
+For the MNIST experiments (i) end-to-end precision, (ii) timing, (iii)
+inference-only precision reduction, and (iv) automatic mixed precision (AMP)
+with gradient clipping, we use 5-fold cross-validation on the training set with
+random equal splits. Unless noted otherwise, figures report fold means, and no
+test set tuning is performed. In the end-to-end runs, every floating point
+tensor appearing inside the training loop (parameters, activations, gradients,
+and optimizer state) is kept at the target precision during training and
+evaluation. The timing experiment reuses these runs to measure mean per-epoch
+training time and mean evaluation time. For inference-only reduction, we first
+train a baseline model in FP64 and then do inference on copies cast to FP32 and
+FP16 without any further optimization (so no \textit{retraining} on the lower
+precision).\\
+
+The AMP experiment is run on MNIST with gradient clipping, comparing an AMP
+setup (FP16 computations with FP32 master weights) against a pure FP32
+baseline. We try different global-norm and per-element value-clipping
+thresholds to check stability and report test accuracy and training/inference
+times. We also run a one-off rounding experiment on FashionMNIST (batch size
+$256$, FP32 training): after training, only the stored weights are rounded to
+$k$ decimal places ($k{=}1\ldots14$) and inference remains FP32. Inputs and
+intermediate computations stay at full precision. We report test accuracy as a
+function of $k$ and note the impact on storage.
+
 \newpage
@@
 y-axis is not starting from zero, so the absolute height differences are
-amplified visually, though the delta is only around $.1\%$. Re-training the
+amplified visually, though the delta is only around $.1\%$. Retraining the
 models and re-evaluating them has also lead to the accuracies being
@@
 \subsection*{Automatic Mixed Precision and Gradient Clipping}
-Automatic Mixed Precision (AMP) accelerates deep learning training by using
-both FP16 and FP32 data types reducing memory usage and increasing
-computational speed. However, FP16's limited dynamic range can cause numerical
-instability, especially during backpropagation. To address this, Gradient
-Clipping is used to cap the magnitude of gradients (either globally or per
-element) preventing extreme updates that could hinder convergence. Together,
-AMP and gradient clipping enable efficient, stable training by balancing
-performance and numerical precision.
+Automatic Mixed Precision (AMP) improves deep learning training by using both
+FP16 and FP32 data types reducing memory usage and increasing computational
+speed. However, FP16's limited dynamic range can cause numerical instability,
+especially during backpropagation. To address this, Gradient Clipping is used
+to cap the magnitude of gradients (either globally or per element) preventing
+extreme updates that could hinder convergence. Together, AMP and gradient
+clipping enable efficient, stable training by balancing performance and
+numerical precision.
 
@@
 	\includegraphics[width=0.85\textwidth]{figures/118_accuracy.png}
-	\caption{Test Accuracy vs.\ number of decimal digits retained in weights.}\label{fig:118_acc}
+	\caption{Test accuracy vs.\ number of decimal digits retained in weights.}\label{fig:118_acc}
 \end{figure}
@@
 accuracy the model gets close to the full precision 88\% accuracy with 32 bit
-floating point. Making and educated guess on distribution such as
+floating point. Making an educated guess on distribution such as
 Figure~\ref{fig:118_2} seems to be a good heuristic for this example. At just 3
@@
 
-% This work confirms that for relatively simple tasks like MNIST classification,
-% significant reductions in parameter precision can be achieved without
-% compromising performance. Our future efforts will extend this analysis to more
-% complex datasets (e.g., CIFAR-10, ImageNet) and deeper architectures, and
-% explore advanced quantization-aware training techniques.
-
 This study demonstrates that neural networks can maintain strong performance

### 2025-11-12 paper/main.tex
weiter

@@
 \usepackage{subcaption}
+\usepackage{bbm}
 
@@
 much. This is especially useful for devices with very limited storage. We also
-try out mixed-precision training, using a mix of Half-Precision Floating Point
-(FP16) and Single-Precision Floating Point (FP32) formats. On top of that, we
-experiment with different types of gradient clipping, like global norm clipping
-and value clipping, to see how they affect training stability and final
-accuracy. \\\\ Even though we use relatively simple network architectures and
-datasets, our results show that neural networks are quite robust to
-low-precision settings. In some cases, we were able to round weights to just
-three decimal digits and still get nearly the same accuracy as the
-full-precision version. These findings show that it is possible to
-significantly reduce the memory and compute needs of neural networks without
-losing much performance. This makes them more practical for use in low-resource
-settings. It also opens the door for more advanced techniques that combine
-quantization, pruning, and precision-aware training to build highly efficient
-models for real-world applications.
+try out mixed-precision training, using a mix of Double-Precision Floating
+Point (FP64), Half-Precision Floating Point (FP16) and Single-Precision
+Floating Point (FP32) formats. On top of that, we experiment with different
+types of gradient clipping, like global norm clipping and value clipping, to
+see how they affect training stability and final accuracy. \\\\ Even though we
+use relatively simple network architectures and datasets, our results show that
+neural networks are quite robust to low-precision settings. In some cases, we
+were able to round weights to just three decimal digits and still get nearly
+the same accuracy as the full-precision version. These findings show that it is
+possible to significantly reduce the memory and compute needs of neural
+networks without losing much performance. This makes them more practical for
+use in low-resource settings. It also opens the door for more advanced
+techniques that combine quantization, pruning, and precision-aware training to
+build highly efficient models for real-world applications.
 
@@
 \begin{equation}\label{eq:activation}
-	a = \omega^\top x + b,\quad \hat{y} = \sigma(a),
+	a=\omega^\top x+b,\quad\hat{y}=\sigma(a)
 \end{equation}
@@
 
-\newpage
 \section{Results}

### 2025-11-17 paper/main.tex
weiter

@@
 A typical neuron in a neural network performs a weighted sum of its inputs,
-followed by an activation function (depicted in~\eqref{eq:activation}):
+followed by an activation function which results in the neuron output $h$
+depicted in~\eqref{eq:activation}. Here, $\omega\in\mathbb{R}^n$ is the weight
+vector, $b \in \mathbb{R}$ is the bias, $x \in \mathbb{R}^n$ is the input, and
+$\sigma$ is a nonlinear activation function (e.g., sigmoid or ReLU).
 
 \begin{equation}\label{eq:activation}
-	a=\omega^\top x+b,\quad\hat{y}=\sigma(a)
+	a=\omega^\top x+b,\quad h=\sigma(a)
 \end{equation}
 
-where $\omega \in \mathbb{R}^n$ is the weight vector, $b \in \mathbb{R}$ is the
-bias, $x \in \mathbb{R}^n$ is the input, and $\sigma$ is a nonlinear activation
-function (e.g., sigmoid or ReLU). Training is conducted via backpropagation,
-where gradients of a loss function $L(\hat{y}, y)$ with respect to the model
-parameters are computed. For example, using the mean squared error:
+\noindent
+In our MLP, hidden layers use the Rectified Linear Unit (ReLU), so
+$\sigma=\mathrm{ReLU}$, while the output layer is linear, producing raw scores
+(also called logits) $z$ later used in~\eqref{eq:ce}. ReLU~\eqref{eq:relu}
+keeps positive activations linear and sets negatives to zero. An additional
+benefit is that the computation is very fast and simple compared to more
+complex activation functions.
 
-\begin{align}\label{eq:mse}
-	L(f(x_1, x_2)) & = {(\sigma(\omega^\top x + b) - y)}^2 \nonumber \\
-	               & = {(\sigma(a) - y)}^2 \nonumber                 \\
-	               & = {(\hat y - y)}^2
-\end{align}
-\begin{alignat}{3}\label{eq:grad-w1}
-	\pdv{L}{\omega_1} & = \pdv{L}{\sigma} &  & \pdv{\sigma}{a}      &  & \pdv{a}{\omega_1} \nonumber \\
-	                  & = 2(\hat y - y)~~ &  & \odv{\sigma(a)}{a}~~ &  & x_1
-\end{alignat}
+\begin{equation}\label{eq:relu}
+	\mathrm{ReLU}(a)=\max(0,a),\qquad\frac{d}{da}\mathrm{ReLU}(a)=\mathbbm{1}\{a>0\}
+\end{equation}
+
+\noindent
+Training of the network is conducted via backpropagation, where gradients of a
+loss function $L$ are computed with respect to the model parameters. For
+multi-class classification with $C=10$ classes, the last layer outputs logits
+$z\in\mathbb{R}^C$. To convert the logits into probabilities, usually softmax
+defined as $p_k=\frac{e^{z_k}}{\sum_{j=1}^C e^{z_j}}$ is used. The probability
+of the true class $y$ is then $p_y$. In our experiments, though, we train
+directly on the logits using a simplified version of the cross-entropy loss
+(the negative log-likelihood of a softmax classifier) implemented in
+Pytorch\footnote{\url{https://pytorch.org/}}:
+
+\begin{equation}\label{eq:ce}
+	L(z,y)=-\log p_y=-z_y+\log\sum_{j=1}^C e^{z_j},
+\end{equation}
+
+whose gradients are
+
+\begin{equation}\label{eq:cegrads}
+	\frac{\partial L}{\partial z_k}=p_k-\mathbbm{1}\{k=y\},\qquad\frac{\partial L}{\partial W}=(p-e_y)h^\top,\qquad\frac{\partial L}{\partial b}=p-e_y,
+\end{equation}
+
+with $W$ and $b$ the weight matrix and bias of the last linear layer, $h$ the
+layer input, and $e_y$ the one-hot vector of the target class. Here,
+$\mathbbm{1}\{k=y\}$ is the indicator: it equals $1$ if $k=y$ and $0$
+otherwise.
 
 Memory and computational requirements during training are substantial, as
-gradients, activations, and parameters must be stored and updated iteratively.
-In contrast, inference requires only a forward pass, making it more amenable to
-optimization via precision reduction.
+weights, gradients, activations, and parameters must be stored and updated
+iteratively. Training and inference can both be optimized using precision
+reduction and related techniques, but inference is more well suited to
+optimizations because it only requires a forward pass.
 

### 2025-11-17 paper/main.tex
weiter

@@
 \section{Artificial Neural Networks}
-
 A typical neuron in a neural network performs a weighted sum of its inputs,
@@
 $\mathbbm{1}\{k=y\}$ is the indicator: it equals $1$ if $k=y$ and $0$
-otherwise.
+otherwise.\\
 
+\noindent
 Memory and computational requirements during training are substantial, as
@@
 \section{Memory Efficiency and Numerical Formats}
-
 To reduce memory usage and computational load, precision-aware techniques such
@@
 and terminology of significand, base, and exponent.
-Tables~\ref{tab:binary-bits} and~\ref{tab:binary-precision} provide an overview
-of the different floating point types and their respective characteristics.
 
 \begin{align}\label{eq:floatingPointFormat}
-	\frac{2469}{200}=12.345=\underbrace{12345}_{\text{significand}}\times\underbrace{10^{\overbrace{-3}^{\text{exponent}}}}_{\text{base}}
+	\frac{2469}{200}=12.345=\underbrace{12345}_{\text{significand}}\times{{\underbrace{10}_{\text{base}}}^{\overbrace{-3}^{\text{exponent}}}}
 \end{align}
+
+\noindent
+Tables~\ref{tab:binary-bits} and~\ref{tab:binary-precision} provide an overview
+of the different floating point types and their respective characteristics. The
+\emph{sign}, \emph{exponent}, and \emph{significand} fields specify how many
+bits are allocated to encode the sign of the number, its scale, and its
+significant digits, respectively. The \emph{exponent bias} is a constant added
+to the true exponent so that it can be stored as an unsigned integer. The
+\emph{precision} (in bits) and corresponding \emph{decimal digits} tell how
+many significant digits can be represented accurately in each format.
 

### 2025-11-18 paper/main.tex
weiter

@@
 
-Prior research in this domain can be broadly classified into two categories:
+Prior research in this domain can be broadly divided into methods that use
+quantization already during training and methods that quantize a pretrained
+model later at deployment time.
 
 \subsubsection*{Quantization During Training}
-Several studies have explored training neural networks with reduced precision
-directly. Notable contributions include:
-\begin{itemize}
-	\item \textbf{Mixed Precision Training}, which leverages FP16 arithmetic for speed while maintaining FP32 master weights~\cite{micikevicius2017mixed}.
-	\item \textbf{DoReFa-Net}, which quantizes weights, activations, and gradients to arbitrary bitwidths~\cite{zhou2018dorefa}.
-	\item \textbf{Q-BERT}, a quantized BERT model that maintains accuracy using only 4-bit weights~\cite{shen2020qbert}.
-\end{itemize}
+Mixed-precision training is an approach where most operations are performed in
+FP16, but a separate high-precision copy of the parameters (often called
+\textit{master weights}) is maintained in FP32~\cite{micikevicius2017mixed}.
+The forward and backward passes use FP16 tensors for faster computation and
+reduced memory usage, but parameter updates are applied to the FP32 master
+weights. The updated FP32 weights are then cast back to FP16 for the next
+forward pass. Keeping master weights in higher precision keeps the loss of
+information smaller and helps the training stability.\\
+
+\noindent
+DoReFa-Net extends this concept by quantizing \textit{weights},
+\textit{activations}, and \textit{gradients} to low-precision
+formats~\cite{zhou2018dorefa}. The method allows using \textit{arbitrary
+	bitwidths}, i.e., the number of bits is not restricted to standard IEEE
+floating-point formats, but can be any small integer (for example, 1, 2, or 4
+bits for weights and slightly higher bitwidths for activations and gradients).
+This is flexible and makes it possible to set the precision specific to
+hardware constraints or application requirements, but the training of the
+network is still end-to-end.\\
+
+\noindent
+Another work is Q-BERT~\cite{shen2020qbert}, that focuses on transformer-based
+language models. BERT (Bidirectional Encoder Representations from Transformers)
+is a large transformer architecture pretrained on large text data and used as a
+generic language representation model. Q-BERT shows that the weights of BERT
+can be quantized down to 4 bits while maintaining accuracy close to the
+full-precision model.
 
 \subsubsection*{Post-Training Quantization}
-Other approaches focus on converting pretrained models into lower-precision
-equivalents:
-\begin{itemize}
-	\item \textbf{Integer-Only Quantization} using symmetric quantization scales~\cite{jacob2018quantization}.
-	\item \textbf{GPTQ}, a method for quantizing large transformer models without retraining~\cite{frantar2022gptq}.
-	\item \textbf{SmoothQuant}, which scales activations to better match weight ranges during quantization~\cite{xiao2024smoothquant}.
-\end{itemize}
+Post-training quantization begins with an already trained full-precision model
+and converts it to a lower precision without (or with minimal) additional
+training. An often used approach is integer-only quantization with symmetric
+scaling~\cite{jacob2018quantization}. Here, both weights and activations are
+mapped to signed integers (for example 8-bit) using a scale factor that is the
+same across positive and negative values (therefore symmetric). This design
+allows the entire inference flow to be implemented with integer arithmetic,
+which is better for embedded or mobile hardware.\\
 
-\section{Artificial Neural Networks}
+\noindent
+Methods such as GPTQ (Gradient Post-Training Quantization) have been tested to
+quantize large transformer models. Examples are GPT language models with
+minimal or no retraining~\cite{frantar2022gptq}. GPTQ treats quantization as an
+optimization problem by using approximations of the local curvature to choose
+quantization levels that minimize the error introduced by replacing
+full-precision weights with low-precision ones. That makes this approach better
+for very large models where a full retraining with quantization would be too
+expensive.\\
+
+\noindent
+SmoothQuant focuses more on activations~\cite{xiao2024smoothquant}. It rescales
+activations and compensates for this rescaling in the weight matrices so that
+the activation distributions become more smooth and better aligned with the
+range supported by low-precision integer formats. This improves the robustness
+of post-training quantization, especially in transformer architectures where
+activation outliers can otherwise complicate the scaling and lead to large
+quantization errors.
+
+\subsection{Artificial Neural Networks}
 A typical neuron in a neural network performs a weighted sum of its inputs,
@@
 
-\section{Memory Efficiency and Numerical Formats}
+\subsection{Memory Efficiency and Numerical Formats}
 To reduce memory usage and computational load, precision-aware techniques such
@@
 of the different floating point types and their respective characteristics. The
-\emph{sign}, \emph{exponent}, and \emph{significand} fields specify how many
+\textit{sign}, \textit{exponent}, and \textit{significand} fields specify how many
 bits are allocated to encode the sign of the number, its scale, and its
-significant digits, respectively. The \emph{exponent bias} is a constant added
+significant digits, respectively. The \textit{exponent bias} is a constant added
 to the true exponent so that it can be stored as an unsigned integer. The
-\emph{precision} (in bits) and corresponding \emph{decimal digits} tell how
+\textit{precision} (in bits) and corresponding \textit{decimal digits} tell how
 many significant digits can be represented accurately in each format.
@@
 	\end{tabular}
-	\caption{Range and precision characteristics of the same formats (adapted from~\cite{fparithmetic}).}\label{tab:binary-precision}
+	\caption{Range and precision characteristics of the formats (adapted from~\cite{fparithmetic}).}\label{tab:binary-precision}
 \end{table}
@@
 \section{Summary}
-% \begin{itemize}
-% 	\item \textbf{16-bit is enough for this classification task}: End-to-end FP16 preserved the accuracy.
-% 	\item \textbf{Three decimals suffice}: Storing weights down to fractional length $3$ caused no visible accuracy loss.
-% \end{itemize}
-
-% \section{Discussion}
-% The results of the first experiments have shown that either the selected task (MNIST
-% classification) was not complex enough to be affected by the decrease in accuracy, or the
-% accuracy would need to be decreased even further than FP16. Otherwise, the results
-% generally indicate that networks trained and evaluated with as low as 16-bit precision can
-% achieve comparable performance to their high-precision counterparts. Moreover, rounding
-% pretrained weights to just 3 decimal digits preserves full accuracy, opening the door to
-% aggressive post-training quantization.This is particularly relevant for edge devices and
-% real-time applications where memory and compute constraints are more relevant.
 

### 2025-11-18 paper/main.tex
weiter

@@
 \subsection{Related Work}
-
 Prior research in this domain can be broadly divided into methods that use

### 2025-11-30 paper/main.tex
weiter

@@
 \usepackage{bbm}
+\usepackage{siunitx}
 
@@
 	neural networks can be reduced without sacrificing too much of the performance.
-	Using MNIST as a general dataset (and FashionMNIST for the decimal rounding),
-	we benchmark uniform quantization (mapping real values to evenly spaced levels
-	via a single step size), mixed-precision formats, and post-training decimal
-	rounding. We evaluate precision settings (FP64/FP32/FP16) across all tensors
-	during training, experiment with an inference-only quantization path on
-	FP64-trained models, and test automatic mixed precision with gradient clipping.
-	Metrics include test accuracy, training/inference time, and parameter storage.
-	We compare end-to-end low-precision training, inference-only quantization, and
-	mixed-precision with gradient clipping, and evaluate their impact on test
-	accuracy and efficiency. The results on the used test setup show that (i)
-	16-bit end-to-end training does not significantly decrease test accuracy for
-	the MNIST dataset, (ii) three-decimal-place weight rounding while keeping
-	32-bit precision incurs minimal accuracy loss. These findings verify limits for
-	memory- and energy-constrained usecases like inference in mobile devices.
+	Using MNIST as a general dataset, and FashionMNIST for the decimal rounding, we
+	benchmark training with reduced precision (FP64/FP32/FP16), mixed-precision
+	formats, and post-training decimal rounding. We evaluate precision settings
+	across all tensors during training, experiment with an inference-only casting
+	to lower-precision floating point on FP64-trained models, and test Automatic
+	Mixed Precision (AMP) with gradient clipping. Metrics include test accuracy,
+	training/inference time, and parameter storage. We compare end-to-end
+	low-precision training, inference-only quantization, and mixed-precision with
+	gradient clipping, and evaluate their impact on test accuracy and efficiency.
+	The results on the used test setup show that (i) 16-bit end-to-end training
+	does not significantly decrease test accuracy for the MNIST dataset, (ii)
+	three-decimal-place weight rounding while keeping 32-bit precision incurs
+	minimal accuracy loss. These findings verify limits for memory- and
+	energy-constrained usecases like inference in mobile devices.
 \end{abstract}
@@
 better results than using a single fixed precision for the whole model.
-Researchers have used a range of techniques, like stochastic gradient descent
+Researchers have used a range of techniques, like Stochastic Gradient Descent
 (SGD), heuristics, and even evolutionary algorithms, to find the best bitwidth
@@
 studies have shown that reducing the precision of neural network parameters can
-lead to significant improvements in memory usage and processing speed. Even
-when precision is reduced to very low levels, models can still perform
-competitively on complex tasks. However, while these early results are
-promising, the overall effects of lowering precision, especially how it
-influences model accuracy, training behavior, and storage efficiency, are still
-not fully understood and deserve more detailed investigation. \\\\ This is
+lead to significant improvements in memory usage and inference efficiency
+without degrading the accuracy too much. For example, integer-only INT8
+quantization~\cite{jacob2018quantization} and also binary-weight and
+binary-activation CNNs evaluated on ImageNet~\cite{rastegariECCV16}. However,
+while these early results are promising, the accuracy/efficiency trade-off
+depends on architecture, task, storage efficiency, and training setup, so
+precision reduction still deserves more detailed investigation. \\\\ This is
 where our experimental work begins. In this paper, we take a closer look at how
@@
 formats~\cite{zhou2018dorefa}. The method allows using \textit{arbitrary
-	bitwidths}, i.e., the number of bits is not restricted to standard IEEE
+	bitwidths} which means the number of bits is not restricted to standard IEEE
 floating-point formats, but can be any small integer (for example, 1, 2, or 4
@@
 and converts it to a lower precision without (or with minimal) additional
-training. An often used approach is integer-only quantization with symmetric
+training. An often used approach is integer quantization with symmetric
 scaling~\cite{jacob2018quantization}. Here, both weights and activations are
@@
 quantize large transformer models. Examples are GPT language models with
-minimal or no retraining~\cite{frantar2022gptq}. GPTQ treats quantization as an
-optimization problem by using approximations of the local curvature to choose
-quantization levels that minimize the error introduced by replacing
+minimal or no retraining~\cite{frantar2022gptq}. GPTQ handles quantization as
+an optimization problem by using approximations of the local curvature to
+choose quantization levels that minimize the error introduced by replacing
 full-precision weights with low-precision ones. That makes this approach better
@@
 the activation distributions become more smooth and better aligned with the
-range supported by low-precision integer formats. This improves the robustness
-of post-training quantization, especially in transformer architectures where
+range supported by low-precision integer formats. This improves the
+post-training quantization, especially in transformer architectures where
 activation outliers can otherwise complicate the scaling and lead to large
@@
 vector, $b \in \mathbb{R}$ is the bias, $x \in \mathbb{R}^n$ is the input, and
-$\sigma$ is a nonlinear activation function (e.g., sigmoid or ReLU).
+$\sigma$ is a nonlinear activation function (sigmoid or ReLU, for example).
 
@@
 keeps positive activations linear and sets negatives to zero. An additional
-benefit is that the computation is very fast and simple compared to more
+benefit is that the calculation is very fast and simple compared to more
 complex activation functions.
@@
 layer input, and $e_y$ the one-hot vector of the target class. Here,
-$\mathbbm{1}\{k=y\}$ is the indicator: it equals $1$ if $k=y$ and $0$
+$\mathbbm{1}\{k=y\}$ is the indicator: it equals \num{1} if $k=y$ and \num{0}
 otherwise.\\
@@
 Memory and computational requirements during training are substantial, as
-weights, gradients, activations, and parameters must be stored and updated
-iteratively. Training and inference can both be optimized using precision
-reduction and related techniques, but inference is more well suited to
-optimizations because it only requires a forward pass.
+weights, gradients, and activations must be stored and updated iteratively.
+Training and inference can both be optimized using precision reduction and
+related techniques, but inference is more well suited to optimizations because
+it only requires a forward pass.
 
@@
 	\url{http://yann.lecun.com/exdb/mnist/}.} ($28\times 28$ grayscale images of
-handwritten digits, train/test split $60.000$/$10.000$, 10 classes, examples
-shown in Figure~\ref{fig:mnistSample}) is used to determine precision effects
-on accuracy and runtime under a simple workload, while an after-training
-rounding experiment is performed on FashionMNIST\footnote{Xiao, Rasul, and
-	Vollgraf: \url{https://github.com/zalandoresearch/fashion-mnist}.} ($28\times
-28$ grayscale images of clothing items, train/test split $60.000$/$10.000$, 10
-classes) so the impact of using coarse value sets is better observable.\\
+handwritten digits, train/test split \num{60000}/\num{10000}, \num{10} classes,
+examples shown in Figure~\ref{fig:mnistSample}) is used to determine precision
+effects on accuracy and runtime under a simple workload, while an
+after-training rounding experiment is performed on FashionMNIST\footnote{Xiao,
+	Rasul, and Vollgraf: \url{https://github.com/zalandoresearch/fashion-mnist}.}
+($28\times 28$ grayscale images of clothing items, train/test split
+\num{60000}/\num{10000}, \num{10} classes) so the impact of using coarse value
+sets is better observable.\\
 
@@
 
-The classifier used is a fully connected network
-$784\rightarrow128\rightarrow64\rightarrow10$ with ReLU between linear layers
-(Figure~\ref{fig:mlpStructure}). Inputs are normalized with mean $0.1307$ and
-std $0.3081$. Unless stated otherwise, we train with cross-entropy loss and SGD
-(learning rate $0.01$, momentum $0.9$), batch size $64$, and on CUDA.\\
+\noindent
+The classifier used is a fully connected network (layers:
+$784\rightarrow128\rightarrow64\rightarrow10$) with ReLU between linear layers
+(see structure in Figure~\ref{fig:mlpStructure}). The final linear layer
+produces \num{10} logits and during training we use \textit{CrossEntropyLoss}
+on the logits, softmax is not applied in the forward pass. Optimization uses
+SGD with learning rate \num{0.01} and momentum \num{0.9}, batch size \num{64},
+on CUDA.\ Unless noted otherwise we train for a small, fixed number of epochs
+per fold. Automatic mixed precision runs use \textit{autocast} with a
+\textit{GradScaler}. When enabled, gradients are clipped either by global norm
+or per-element value as specified in Section~\ref{sec:results}.
 
@@
 
+\noindent
 For the MNIST experiments (i) end-to-end precision, (ii) timing, (iii)
-inference-only precision reduction, and (iv) automatic mixed precision (AMP)
-with gradient clipping, we use 5-fold cross-validation on the training set with
-random equal splits. Unless noted otherwise, figures report fold means, and no
-test set tuning is performed. In the end-to-end runs, every floating point
-tensor appearing inside the training loop (parameters, activations, gradients,
-and optimizer state) is kept at the target precision during training and
-evaluation. The timing experiment reuses these runs to measure mean per-epoch
-training time and mean evaluation time. For inference-only reduction, we first
-train a baseline model in FP64 and then do inference on copies cast to FP32 and
-FP16 without any further optimization (so no \textit{retraining} on the lower
-precision).\\
+inference-only precision reduction, and (iv) AMP with gradient clipping, we use
+5-fold cross-validation on the training set with random equal splits. Unless
+noted otherwise, figures report fold means, and no test set tuning is
+performed. In the end-to-end runs, every floating point tensor appearing inside
+the training loop (parameters, activations, gradients) is kept at the target
+precision during training and evaluation. The timing experiment reuses these
+runs to measure mean per-epoch training time and mean evaluation time. For
+inference-only reduction, we first train a baseline model in FP64 and then do
+inference on copies cast to FP32 and FP16 without any further optimization (so
+no \textit{retraining} on the lower precision).\\
 
@@
 times. We also run a one-off rounding experiment on FashionMNIST (batch size
-$256$, FP32 training): after training, only the stored weights are rounded to
-$k$ decimal places ($k{=}1\ldots14$) and inference remains FP32. Inputs and
+\num{256}, FP32 training): after training, only the stored weights are rounded
+to $k$ decimal places ($k{=}1\ldots14$) and inference remains FP32. Inputs and
 intermediate computations stay at full precision. We report test accuracy as a
@@
 
-\section{Results}
+\section{Experiments and Results}\label{sec:results}
 In the first experiment, every floating-point tensor that is used inside the
@@
 constrains each individual gradient component within a fixed range, directly
-limiting extreme values. These methods are evaluated under Automatic Mixed
-Precision (AMP) to compare their impact on training stability, model accuracy,
-and computation time across different precision types.
+limiting extreme values. These methods are evaluated under AMP to compare their
+impact on training stability, model accuracy, and computation time across
+different precision types.
 

### 2025-11-30 paper/main.tex
weiter

@@
 \noindent
-For the MNIST experiments (i) end-to-end precision, (ii) timing, (iii)
-inference-only precision reduction, and (iv) AMP with gradient clipping, we use
-5-fold cross-validation on the training set with random equal splits. Unless
-noted otherwise, figures report fold means, and no test set tuning is
+For the MNIST experiments (i) FP64/FP32/FP16 end-to-end training, (ii) effects
+on timing, (iii) inference-only casting, and (iv) AMP with gradient clipping,
+we use 5-fold cross-validation on the training set with random equal splits.
+Unless noted otherwise, figures report fold means, and no test set tuning is
 performed. In the end-to-end runs, every floating point tensor appearing inside
@@
 runs to measure mean per-epoch training time and mean evaluation time. For
-inference-only reduction, we first train a baseline model in FP64 and then do
+inference-only casting, we first train a baseline model in FP64 and then do
 inference on copies cast to FP32 and FP16 without any further optimization (so
-no \textit{retraining} on the lower precision).\\
+no \textit{retraining} on the lower precision). The AMP experiment is run on
+MNIST with gradient clipping, comparing an AMP setup (FP16 computations with
+FP32 master weights) against a pure FP32 baseline. We try different global-norm
+and per-element value-clipping thresholds to check stability and report test
+accuracy and training/inference times.\\
 
-The AMP experiment is run on MNIST with gradient clipping, comparing an AMP
-setup (FP16 computations with FP32 master weights) against a pure FP32
-baseline. We try different global-norm and per-element value-clipping
-thresholds to check stability and report test accuracy and training/inference
-times. We also run a one-off rounding experiment on FashionMNIST (batch size
-\num{256}, FP32 training): after training, only the stored weights are rounded
-to $k$ decimal places ($k{=}1\ldots14$) and inference remains FP32. Inputs and
-intermediate computations stay at full precision. We report test accuracy as a
-function of $k$ and note the impact on storage.
+\noindent
+On the FashionMNIST dataset, experiment (v) is a rounding experiment (batch
+size \num{256}, FP32 training) which means that after training only the stored
+weights are rounded to $k$ decimal places ($k{=}1\ldots14$) and inference
+remains at FP32. Inputs and intermediate computations stay at full precision.
+We report test accuracy as a function of $k$ and note the impact on storage.
 
@@
 \subsection*{Automatic Mixed Precision and Gradient Clipping}
-Automatic Mixed Precision (AMP) improves deep learning training by using both
-FP16 and FP32 data types reducing memory usage and increasing computational
-speed. However, FP16's limited dynamic range can cause numerical instability,
-especially during backpropagation. To address this, Gradient Clipping is used
-to cap the magnitude of gradients (either globally or per element) preventing
-extreme updates that could hinder convergence. Together, AMP and gradient
-clipping enable efficient, stable training by balancing performance and
-numerical precision.
+AMP improves deep learning training by using both FP16 and FP32 data types
+reducing memory usage and increasing computational speed. However, FP16's
+limited dynamic range can cause numerical instability, especially during
+backpropagation. To address this, Gradient Clipping is used to cap the
+magnitude of gradients (either globally or per element) preventing extreme
+updates that could hinder convergence. Together, AMP and gradient clipping
+enable efficient, stable training by balancing performance and numerical
+precision.
 

### 2025-11-30 paper/main.tex
fertig

@@
 \usepackage{siunitx}
+\usepackage{refcount}
 
@@
 can be quantized down to 4 bits while maintaining accuracy close to the
-full-precision model.
+full-precision model.\\
+
+\noindent
+AMP improves deep learning training by using both FP16 and FP32 data types
+reducing memory usage and increasing computational speed. However, FP16's
+limited dynamic range can cause numerical instability, especially during
+backpropagation. To address this, Gradient Clipping is used to cap the
+magnitude of gradients (either globally or per element) preventing extreme
+updates that could hinder convergence. Together, AMP and gradient clipping
+enable efficient, stable training by balancing performance and numerical
+precision~\cite{micikevicius2017mixed}.
 
@@
 In our MLP, hidden layers use the Rectified Linear Unit (ReLU), so
-$\sigma=\mathrm{ReLU}$, while the output layer is linear, producing raw scores
+$\sigma:=\mathrm{ReLU}$, while the output layer is linear, producing raw scores
 (also called logits) $z$ later used in~\eqref{eq:ce}. ReLU~\eqref{eq:relu}
@@
 \begin{equation}\label{eq:cegrads}
-	\frac{\partial L}{\partial z_k}=p_k-\mathbbm{1}\{k=y\},\qquad\frac{\partial L}{\partial W}=(p-e_y)h^\top,\qquad\frac{\partial L}{\partial b}=p-e_y,
+	\frac{\partial L}{\partial z_k}=p_k-\mathbbm{1}_{k=y},\qquad\frac{\partial L}{\partial W}=(p-e_y)h^\top,\qquad\frac{\partial L}{\partial b}=p-e_y,
 \end{equation}
@@
 layer input, and $e_y$ the one-hot vector of the target class. Here,
-$\mathbbm{1}\{k=y\}$ is the indicator: it equals \num{1} if $k=y$ and \num{0}
-otherwise.\\
+$\mathbbm{1}_{k=y}$ is the indicator: it equals \num{1} if $k=y$ and \num{0}
+otherwise. For evaluation, the model is fed multiple inputs and the results of
+the indicator function are summed up and divided by the total number of inputs.
+This is then called the accuracy score of the model which can be understood as
+the ratio of correct predictions.\\
 
@@
 	\end{tabular}
-	\caption{Field widths (in bits) of common IEEE754 formats (adapted from~\cite{fparithmetic}).}\label{tab:binary-bits}
+	\caption{Field widths (in bits) of common IEEE754 formats\protect\footnotemark.}\label{tab:binary-bits}
 \end{table}
+\footnotetext{\label{footnoteWiki}\url{https://en.wikipedia.org/wiki/Floating-point_arithmetic}}
 
@@
 	\end{tabular}
-	\caption{Range and precision characteristics of the formats (adapted from~\cite{fparithmetic}).}\label{tab:binary-precision}
+	\caption{Range and precision characteristics of the formats\protect\footnotemark[\getrefnumber{footnoteWiki}].}\label{tab:binary-precision}
 \end{table}
@@
 \section{Experiments and Results}\label{sec:results}
-In the first experiment, every floating-point tensor that is used inside the
+In the first experiment, every floating-point variable that is used inside the
 training loop was converted to a lower precision data type. This therefore
-includes the weights and biases, activations, gradients, optimizer states, etc.
+includes the weights and biases, activations, gradients, and optimizer states.
 The types tried were FP64, FP32, and FP16.
@@
 	\includegraphics[width=0.85\textwidth]{figures/trainingAndInference.png}
-	\caption{Model accuracy evaluated after training with different precisions.}\label{fig:modelAccTrainAndInf}
+	\caption{Model accuracy evaluated after training with different precisions (5-fold CV).}\label{fig:modelAccTrainAndInf}
 \end{figure}
@@
 	\includegraphics[width=0.85\textwidth]{figures/times.png}
-	\caption{Training and evaluation times for the different precisions.}\label{fig:modelEvalTrainTimes}
+	\caption{Training and evaluation times for the different precisions (5-fold CV).}\label{fig:modelEvalTrainTimes}
 \end{figure}
@@
 \subsection*{Automatic Mixed Precision and Gradient Clipping}
-AMP improves deep learning training by using both FP16 and FP32 data types
-reducing memory usage and increasing computational speed. However, FP16's
-limited dynamic range can cause numerical instability, especially during
-backpropagation. To address this, Gradient Clipping is used to cap the
-magnitude of gradients (either globally or per element) preventing extreme
-updates that could hinder convergence. Together, AMP and gradient clipping
-enable efficient, stable training by balancing performance and numerical
-precision.
+This experiment uses AMP with different gradient clipping methods. The first
+method is called global norm clipping where all gradients are collectively
+scaled to ensure their combined norm does not exceed a specified threshold,
+which ensures stable updates across the network. The second method is
+per-element clipping which constrains each individual gradient component within
+a fixed range, directly limiting extreme values. These methods are evaluated
+under AMP to compare their impact on training stability, model accuracy, and
+computation time across different precision types.\\
 
-First step is using global norm clipping, all gradients are collectively scaled
-to ensure their combined norm does not exceed a specified threshold, promoting
-stable updates across the network. In second step per-element clipping
-constrains each individual gradient component within a fixed range, directly
-limiting extreme values. These methods are evaluated under AMP to compare their
-impact on training stability, model accuracy, and computation time across
-different precision types.
+\noindent
+Figure~\ref{fig:ampGradientNormClipping} depicts the effect of the
+\textit{global} gradient norm on accuracy, training and testing timings.
+Accuracy is virtually the same between FP16 and FP32 when gradient clipping is
+used validating the stability of AMP and gradient scaling. Eval time is nearly
+identical for both formats, as evaluation involves fewer operations and less
+intense memory bandwidth usage.
 
@@
 	\includegraphics[width=0.95\textwidth]{figures/FP16_32_clipping.png}
-	\caption{Affect of Gradient Norm on accuracy, training and testing timings.
-		Accuracy is virtually the same between FP16 and FP32 when gradient clipping is used
-		validating the stability of AMP and gradient scaling.
-		Eval time is nearly identical for both formats, as evaluation involves fewer operations
-		and less intense memory bandwidth usage.}
+	\caption{Effect of global gradient norm clipping on accuracy, training and testing timings.}\label{fig:ampGradientNormClipping}
 \end{figure}
@@
 	\includegraphics[width=0.95\textwidth]{figures/gradclip_16.png}
-	\caption{Effect of Per-Element Gradient Clipping on FP16. Accuracy vs Training \& Eval Time}\label{fig:grad16}
+	\caption{Effect of per-element gradient clipping on FP16 $\rightarrow$ Accuracy vs training and evaluation time}\label{fig:grad16}
 \end{figure}
@@
 	\includegraphics[width=0.95\textwidth]{figures/gradclip_32.png}
-	\caption{Effect of Per-Element Gradient Clipping on FP32. Accuracy vs Training \& Eval Time}\label{fig:grad32}
+	\caption{Effect of per-element gradient clipping on FP32 $\rightarrow$ Accuracy vs training and evaluation time}\label{fig:grad32}
 \end{figure}
 
-Figure~\ref{fig:grad16} shows accuracy peaked at 97.07\% with a clipping value
-of 0.5, while the lowest accuracy (96.55\%) occurred at clip=0.01, suggesting
-that very small clip values may overly restrict gradient updates and hurt
-learning. Training time ranged indicating that moderate clipping supports
-efficient convergence. Evaluation time remained stable across all clip values
-(around 2.01s to 2.16s), showing that clipping primarily impacts training
-rather than inference performance. Also Figure~\ref{fig:grad32} shows that
+\noindent
+Figure~\ref{fig:grad16} shows that accuracy peaked at 97.07\% with a clipping
+value of 0.5, while the lowest accuracy (96.55\%) occurred at clip=0.01,
+suggesting that very small clip values may overly restrict gradient updates and
+hurt learning. Training time varied, which indicates that moderate clipping
+supports efficient convergence. The evaluation time remained stable across all
+clip values (around 2.01s to 2.16s), showing that clipping primarily impacts
+training rather than inference performance. Figure~\ref{fig:grad32} shows that
 accuracy was most stable across all clip values with the highest accuracy at
-clip=0.5, demonstrating its robustness to gradient clipping variations.
-Training time suggesting that larger clip values can speed up convergence
-without sacrificing stability. Evaluation time confirming that clipping had
-minimal effect on inference time.
+clip=0.5, demonstrating its robustness to gradient clipping variations. The
+training was conducted with 5-fold CV, 10 epochs, and batch size 64. The
+training time suggests that larger clip values can speed up convergence without
+sacrificing stability. Evaluation time confirms that clipping had only a minor
+effect on inference time.
 
 \subsection*{Effect of Decimal Rounding}
-
 A non training related issue is the storage of the NN parameters. Usually, the
@@
 model weights. The datatype precision has a higher impact during training
-though, which begs the question if one could omit precision from the model
+though, which begs the question if one could reduce the precision of the model
 weights. In this section, we explore a naive quantization method: Rounding to a
-fixed number of decimals. With significantly little decimals and a reasonable
-range for the integer part, one could store only the significant and provide
-base as well as exponent as metadata. When loading the model, the weights are
-converted back to floating point, with whatever precision is desired. Only the
-weights are quantized; the input data, as well are the values propagated
-through the network are at full precision.
+fixed number of decimals. With a small enough number of decimals and a
+reasonable range for the integer part, one could store only the significant and
+provide base as well as exponent as separate data. When loading the model, the
+weights are converted back to floating point, with whatever precision is
+desired. Only the weights are quantized; the input data, as well are the values
+propagated through the network are at full precision.\\
 
+\noindent
 The first step is to examine the distribution of parameter values. By the
@@
 Reading from the distribution is just a heuristic though, so we will also test
-the hypotheses.
+the hypotheses.\\
+
+\noindent
+Figure~\ref{fig:roundingDistribution1} shows the distribution of the model
+weights in a three layer architecture. The values are in $[-0.5, 0.5]$, which
+means no integer part is needed in this example. In Figure~\ref{fig:118_2} the
+weights of the same model rounded to two decimals is depicted. The envelope of
+the bins is still very similar to the full precision distribution, which makes
+it appear like a good approximation. The results of evaluating this heuristic
+is shown in Figure 11, where the results of how rounding the decimal part
+affects the test accuracy is shown.
 
@@
 	\caption{
-		Distribution of model weights in each layer. The values are in $[-0.5, 0.5]$,
-		this means no integer part is needed in this example. Generally, with normalized data
-		and the quite popular L1-/L2-/Batch-Norm layers, weight values in $[-1, 1]$ is to be expected.}
+		Distribution of model weights in each layer.}\label{fig:roundingDistribution1}
 \end{figure}
@@
 	\includegraphics[width=0.85\textwidth]{figures/118_dist2.png}
-	\caption{Distribution of model weights in each layer after rounding to two decimals.
-		This appears to be a good approximation already.}\label{fig:118_2}
+	\caption{Distribution of model weights in each layer after rounding to two decimals.}\label{fig:118_2}
 \end{figure}
@@
 
+\noindent
 Figure~\ref{fig:118_acc} shows how rounding the decimal part affects the test
@@
 with a 10 class problem. Two decimals perform significantly better. At 70\%
-accuracy the model gets close to the full precision 88\% accuracy with 32 bit
-floating point. Making an educated guess on distribution such as
+accuracy the model gets close to the full precision 88\% accuracy with 32 Bit
+floating point. Making an educated guess the on distribution such as
 Figure~\ref{fig:118_2} seems to be a good heuristic for this example. At just 3
 decimals, the full 32 bit precision is reached. Thus we can conclude that with
-just 10 bit for representing 3 decimals, no bits for the integer part and one
-additional bit for the sign, only 11 bit must be stored to have the best
-performing model for inference.
+just \num{10} Bits for representing 3 decimals, no bits for the integer part
+and one additional bit for the sign, only \num{11} Bits must be stored to have
+the best performing model for inference. The storage size of the model is
+reduced to $\frac{11}{32}\approx 34\%$ of the otherwise full \num{32} Bit
+model.
 
-\section{Summary}
+\section{Summary and Conclusion}
 
-The experimental results presented in this paper demonstrate the robustness of
-neural networks to reduced parameter precision, particularly for simpler tasks
-like MNIST classification. Key findings include:
+Regarding the precision reduction during training and inference, models with
+16-bit floating-point (FP16) precision preserved accuracy comparable to
+higher-precision formats (FP32 and FP64). This suggests that for tasks of
+similar complexity, FP16 is sufficient, offering potential memory and
+computational savings without sacrificing performance. Reducing precision only
+during inference, after training with high-precision (FP64), maintained
+baseline accuracy. This is particularly valuable for deployment scenarios where
+training is done on powerful hardware, but inference occurs on
+resource-constrained devices. Training with mixed precision and Gradient
+Clipping resulted in similar test accuracy scores as precision reduction during
+training as well as post-training quantization. No significant decrease in
+model performance accuracy was detected. The train and test times were similar,
+regardless of using FP32, FP16 or a mix of both. Rounding trained weights to
+just three decimal digits (approximately 11 bits) achieved full accuracy,
+highlighting the potential for aggressive quantization in storage-efficient
+model deployment. This finding is significant for edge applications where
+memory footprint is critical.\\
 
-\begin{itemize}
-	\item \textbf{Precision Reduction During Training and Inference}: Training and
-	      evaluating models with 16-bit floating-point (FP16) precision preserved accuracy
-	      comparable to higher-precision formats (FP32 and FP64). This suggests that for
-	      tasks of similar complexity, FP16 is sufficient, offering potential memory and
-	      computational savings without sacrificing performance.
-
-	\item \textbf{Post-Training Quantization}: Reducing precision only during inference, after
-	      training with high-precision (FP64), maintained baseline accuracy. This is
-	      particularly valuable for deployment scenarios where training is done on powerful
-	      hardware, but inference occurs on resource-constrained devices.
-
-	\item \textbf{Mixed-Precision:} Training with mixed precision and Gradient Clipping
-	      resulted in similar test accuracy scores as precision reduction during training
-	      as well as post-training quantization. No significant decrease in model
-	      performance accuracy was detected. The train and test times were similar,
-	      regardless of using FP32, FP16 or a mix of both.
-
-	\item \textbf{Weight Rounding:} Rounding trained weights to just three decimal digits
-	      (approximately 11 bits) achieved full accuracy, highlighting the potential for
-	      aggressive quantization in storage-efficient model deployment. This finding is
-	      significant for edge applications where memory footprint is critical.
-\end{itemize}
-
+\noindent
 However, the simplicity of the MNIST dataset and the small network architecture
@@
 flow and convergence, was not deeply explored and could be a focus for future
-work.
+work.\\
 
-\section{Conclusion}
-
+\noindent
 This study demonstrates that neural networks can maintain strong performance
@@
 models in resource-constrained environments, such as edge devices and real-time
-applications.
+applications.\\
 
+\noindent
 To build on these insights and expand their applicability, future work should
